# Build a `Fleet` Object {#sec-fleet_build}

```{r}
#| eval: true
#| include: false
#| message: false
#| warning: false
#| cache: false
library(MSEtool)
knitr::read_chunk('../MSEtool/data-raw/ExampleOM.R')
```

Each `Stock` in an `OM` must have a corresponding `Fleet` object that describes the exploitation pattern, including fishing effort, the selectivity and retention schedules, and discard mortality.

In most cases, the contents of the `Fleet` object(s) will be populated automatically by the `Condition` (@sec-om_condition) or `Import` (@sec-om_import) functions.

In this example, a `Fleet` object is created using the same expert judgment approach used for the `Stock` object in the previous chapter. 

## Create a new `Fleet` Object

The `Fleet()` function is used to create new `Fleet` objects:

```{r create-fleet}

```


In addition to `Name` that was specified in the call to the `Fleet()` function, `Fleet` objects contain several other sub-objects that need to be populated:

1.  [`Effort`](object_effort.html)
2.  [`Catchability`](object_catchability.html)
3.  [`Weight`](object_weight.html)
4.  [`NaturalMortality`](object_naturalmortality.html)
5.  [`Maturity`](object_maturity.html)
6.  [*`Fecundity`*](object_fecundity.html)
7.  [`SRR`](object_srr.html)
8.  [*`Spatial`*](object_spatial.html)
9.  [*`Depletion`*](object_depletion.html)

The sub-objects in italics are optional, see the relevant chapters for each object for details. The other sub-objects must be populated to complete the `Stock` object.


At the minimum, a `Fleet` object must contain the following information:


The rest of contents of the `Fleet` object could also be added directly in the call to `Fleet` (see `?Fleet`). However for the purposes of demonstration, these contents are added in separate steps in the sections below.



## Fishing `Effort`




