# Reference Points {#sec-refpoints}

```{r}
#| warning: false
#| message: false
#| echo: false
#| include: false
library(openMSE)
```

Reference points are calculated from the operating model during `Simulate`, are stored in `Hist@Reference`, and can be accessed with `Reference(Hist)`.

Three broad categories of reference points are used:

**1. Unfished reference points:** quantities under no fishing, including unfished biomass ($B_0$), unfished spawning biomass ($SB_0$), unfished spawning production ($SP_0$), and unfished recruitment ($R_0$). These describe the baseline state of the stock and are calculated as part of the unfished equilibrium (see @sec-tech-unfished).

**2. MSY-based reference points:** quantities at or derived from maximum sustainable yield, including $F_\text{MSY}$, $B_\text{MSY}$, $SB_\text{MSY}$, and $SP_\text{MSY}$ (see @sec-tech-refpoints-msy).

**3. Per-recruit and replacement-line reference points:** further $F$-based benchmarks derived from the yield-per-recruit curve and the stock-recruitment replacement line, including $F_{0.1}$, $F_\text{Max}$ (see @sec-tech-refpoints-ypr), and $F_{x\%\text{SPR}}$, $F_\text{Med}$, $F_\text{crash}$, and $\text{SPR}_\text{crash}$ (see @sec-tech-refpoints-spr).

Additional reference points include the mean generation time (MGT), the reference yield under a fixed-$F$ policy, and $B_\text{Low}$ (a rebuilding-based limit reference point used in DFO Canada's Precautionary Approach framework).

## MSY Reference Points

MSY reference points are calculated automatically by `Simulate()` and stored in `Hist@Reference@MSY`. They can also be recalculated at any time using `CalcMSY()`, which takes a `Hist` or `OM` object and returns a `refpointsMSY` object.

```{r}
#| eval: false
Hist <- Simulate(SingleStockOM, silent = TRUE)

# Full reference object
Reference(Hist)

# Full refpointsMSY object
MSYRefs(Hist)

# Individual quantities
FMSY(Hist)
BMSY(Hist)
SBMSY(Hist)
MSYLandings(Hist)
```

By default, reference points are evaluated using parameters from the last historical year. Alternative years can be specified via the `Years` argument:

```{r}
#| eval: false
# Reference points at a specific year
CalcMSY(Hist, Years = 2010)

# Reference points at multiple years (evaluated independently)
CalcMSY(Hist, Years = c(2005, 2010, 2020))
```

Evaluating reference points at multiple years can be useful for examining how $F_\text{MSY}$ or $B_\text{MSY}$ shift as biological or fishery parameters change over time.

`CalcMSY()` returns a `refpointsMSY` object with the following slots, all with dimensions `Sim × Stock × Year`:

| Slot | Description |
|------|-------------|
| `FMSY` | Apical fishing mortality at MSY |
| `BMSY` | Total biomass at MSY |
| `SBMSY` | Spawning biomass at MSY |
| `SPMSY` | Spawning production at MSY |
| `SPRMSY` | Spawning potential ratio at MSY |
| `MSYLandings` | Landed catch at MSY |
| `MSYDiscards` | Dead discards at MSY |

### Removals vs Landings

`CalcMSY()` can maximise either total removals (landings plus dead discards) or landings alone, controlled by the `type` argument:

```{r}
#| eval: false
CalcMSY(Hist, type = "Removals")  # default
CalcMSY(Hist, type = "Landings")
```

Removals-based MSY finds the $F$ that maximises total fish removed from the population while landings-based MSY finds the $F$ that maximises retained catch. 

### Multi-Stock Complexes

When a stock complex contains more than one stock, `CalcMSY()` finds the single $F_\text{MSY}$ maximising total yield across all stocks. Stock-level quantities (BMSY, SBMSY, etc.) are then evaluated at this complex-level $F_\text{MSY}$. See @sec-tech-refpoints for details.

## Per-Recruit and Replacement-Line Reference Points

Six further $F$-based reference points are calculated automatically by `Simulate()` and stored in `Hist@Reference`. They can also be recalculated at any time using `CalcRefPoints()`, which takes a `Hist` or `OM` object:

```{r}
#| eval: false
Hist <- CalcRefPoints(Hist)

# Individual quantities
F01(Hist)
FMax(Hist)
FSPR(Hist)
FMed(Hist)
FCrash(Hist)
SPRcrash(Hist)
```

| Quantity | Description |
|------|-------------|
| `F01` | Apical $F$ at which the slope of the yield-per-recruit (YPR) curve is 10% of its slope at the origin |
| `FMax` | Apical $F$ maximising YPR.  |
| `FSPR` | Apical $F$ at which spawning-per-recruit relative to unfished equals a user-specified target (`SPRTarget`, e.g. `0.4` for $F_{40\%}$). Has an additional `Target` dimension for multiple targets, default `c(0.3, 0.4, 0.5)` |
| `FMed` | Apical $F$ at which the per-recruit replacement line equals the median historical recruits-per-spawner ratio |
| `FCrash` | Apical $F$ at which the per-recruit replacement line crosses the compensation ratio implied by steepness, i.e. the $F$ beyond which the stock can no longer replace itself.  |
| `SPRcrash` | The spawning potential ratio at `FCrash` |

`F01` and `FMax` are derived from the yield-per-recruit curve alone and don't require a stock-recruitment relationship. `FMed` and `FCrash`/`SPRcrash` compare the per-recruit replacement line against, respectively, the median historical recruits-per-spawner ratio and the compensation ratio implied by stock-recruitment steepness.

## Reference Yield

The reference yield provides a practical benchmark for comparing MP performance. It is the highest yield (landings or removals) achievable under a fixed-$F$ policy over the projection period, accounting for the actual simulated population trajectory and recruitment variability. 

It is computed separately for landings and removals, stored in `Hist@Reference@Landings` and `Hist@Reference@Removals`. Neither is calculated by default; enable them via `SimControl(RefLandings = TRUE)` and/or `SimControl(RefRemovals = TRUE)` passed to `Simulate()`.

For seasonal models, the effort pattern from the last complete historical year (all $S$ seasons) is extracted as the baseline. A single scalar multiplier is then optimised to scale all fleets and all seasons simultaneously, preserving the relative seasonal effort distribution while searching over the overall level of fishing pressure.

## Mean Generation Time

The mean generation time (MGT) is the mean age of a mature female in the population at unfished equilibrium:

$$\text{MGT} = \frac{\sum_a a \cdot l_a \cdot m_a}{\sum_a l_a \cdot m_a}$$

where $l_a$ is survivorship to age $a$ under natural mortality only and $m_a$ is maturity-at-age. MGT is used as the time unit for the $B_\text{Low}$ calculation below. It is calculated automatically during `Simulate()`, and can be recalculated at any time using `CalcMGT()`:

```{r}
#| eval: false
CalcMGT(Hist)

# At a specific year
CalcMGT(Hist, Years = 2010)
```

MGT is stored in `Hist@Reference@MGT` and has dimensions `Sim × Stock × Year`. 

## B~Low~

$B_\text{Low}$ is a rebuilding-based limit reference point used in DFO Canada's Precautionary Approach framework. It is defined as the current spawning biomass below which a population, if it were to recover with zero fishing, could not rebuild to $B_\text{frac} \times SB_\text{MSY}$ within $\text{HZN} \times \text{MGT}$ years. The default values are $B_\text{frac} = 0.5$ and $\text{HZN} = 2$.

Conceptually, if current spawning biomass falls below $B_\text{Low}$, then even with immediate cessation of fishing the stock would require more than two mean generation times to recover to half of $SB_\text{MSY}$.

$B_\text{Low}$ is stored in `Hist@Reference@BLow`. For each simulation and stock, numbers-at-age in the last historical year are scaled by a factor found by numerical search so that projecting forward under zero fishing effort for $\text{HZN} \times \text{MGT}$ years reaches exactly $B_\text{frac} \times SB_\text{MSY}$; this scaling factor applied to the current spawning biomass gives $B_\text{Low}$.

$B_\text{Low}$ is not calculated by default due to its computational cost. Enable it via `SimControl(BLow = TRUE)` passed to `Simulate()`, or compute it directly with `CalcBLow()`

## Per-Recruit Curves

The underlying per-recruit calculations used to derive $F_\text{MSY}$, $F_{0.1}$, $F_\text{Max}$, and $F_\text{crash}$ can be examined directly using `CalcPerRecruit()` and `CalcEquilibrium()`. These are useful for understanding yield curves, plotting SPR against $F$, and diagnosing model behaviour.

```{r}
#| eval: false
PR <- CalcPerRecruit(Hist, apicalF = seq(0, 1, by = 0.05))
Eq <- CalcEquilibrium(Hist)
```

Both functions accept a `Years` argument and a `Complex` argument.

## Depletion and Stock Status

Current stock status is typically expressed as depletion, the ratio of current biomass to a reference level:

- $B/B_0$: current total biomass relative to unfished biomass
- $SB/SB_0$: current spawning biomass relative to unfished spawning biomass
- $SP/SP_0$: current spawning production relative to unfished spawning production
- $B/B_\text{MSY}$: current biomass relative to $B_\text{MSY}$
- $F/F_\text{MSY}$: current fishing mortality relative to $F_\text{MSY}$

Helper functions for extracting these quantities from a `Hist` or `MSE` object:

```{r}
#| eval: false
# Relative to unfished
B_B0(Hist)       # total biomass / B0
SB_SB0(Hist)     # spawning biomass / SB0
SP_SP0(Hist)     # spawning production / SP0

# Relative to MSY
B_BMSY(Hist)     # total biomass / BMSY
SB_SBMSY(Hist)   # spawning biomass / SBMSY
SP_SPMSY(Hist)   # spawning production / SPMSY
F_FMSY(Hist)     # total apical F (summed over fleets) / FMSY
```

## Seasonal Models {#sec-concept-refpoints-seasonal}

Seasonal operating models subdivide each calendar year into $S$ seasons with season-specific biology, mortality, and recruitment. Reference points for seasonal models are expressed on an annual scale and follow the same logic as annual models. The `Years` argument to `CalcMSY()`, `CalcRefPoints()`, `CalcPerRecruit()`, and `CalcEquilibrium()` should specify complete calendar years (integers); all seasons within each requested year are used automatically.

### Annual reference points from seasonal dynamics

All reference point quantities are reported as annual values. $B_\text{MSY}$, $SB_\text{MSY}$, and $SP_\text{MSY}$ represent equilibrium population state at $F_\text{MSY}$; MSY is total annual catch across all seasons. The same applies to the per-recruit and replacement-line reference points ($F_{0.1}$, $F_\text{Max}$, $F_{x\%\text{SPR}}$, $F_\text{crash}$, $\text{SPR}_\text{crash}$), since they're all derived from the same annually-aggregated per-recruit curve as $F_\text{MSY}$.

In a seasonal model, biomass fluctuates within the year as seasonal mortality and recruitment pulses occur and due to growth within the year. The per-recruit framework accounts for this by computing each quantity as a recruitment-weighted average over birth-season cohorts (see @sec-tech-refpoints-seasonal for the full equations).

### Cohorts and birth seasons

The seasonal recruitment pattern is controlled by `R0` in the `SRR` object: the proportion of annual recruitment entering in season $s_0$ is $\pi_{s_0} = R_{0,s_0} / R_{0,\text{annual}}$. Per-recruit quantities (spawning production, biomass, yield) are computed for each birth season separately and combined using these $\pi_{s_0}$ weights. Full equations are given in @sec-tech-refpoints-seasonal.

### Fecundity and spawning production

`AdjustSeasonalFecundity()` (called automatically during `Simulate()`) scales fecundity-at-age by season in proportion to the seasonal recruitment pattern. This means $\Phi_0$ (unfished spawning production per recruit, the sum of fecundity-weighted survival over all ages and seasons) reflects the actual timing and magnitude of spawning across the year, consistent with whatever recruitment seasonality is specified. SPR $= \Phi_F / \Phi_0$ retains its standard interpretation: the fraction of unfished spawning production per recruit remaining at fishing mortality $F$, regardless of whether spawning is concentrated in one season or distributed across several.

### Interpretation of $F_\text{MSY}$

$F_\text{MSY}$ is the annual apical fishing mortality, the sum of instantaneous rates experienced by the most heavily fished age class across all seasons in a year. This is directly comparable to $F_\text{MSY}$ from an annual model. 

The optimisation scales a peak-season effort scalar, with all other seasons proportional to it via the historical effort pattern; $F_\text{MSY}$ is reported as the corresponding annual total. $F_{0.1}$, $F_\text{Max}$, $F_{x\%\text{SPR}}$, and $F_\text{crash}$ are all reported on this same annual apical-$F$ scale, since they're evaluated over the same grid of annual apical $F$ values.

::: {.callout-note}
`FMed()` is not currently reliable for seasonal models. It compares the per-recruit replacement line (an annually-aggregated quantity) against the median of the raw historical ratio of recruits to spawning biomass at each season-step, rather than an equivalent annual ratio; the two are on different scales, and `FMed` typically comes out `NA`. 
:::

## Spatial Models {#sec-concept-refpoints-spatial}

For spatially structured models, reference points are currently calculated using spatially-aggregated parameters and should be interpreted as spatial averages. A spatially-explicit equilibrium approach is planned as a future option.

For full technical details of all reference point calculations, see @sec-tech-refpoints.
