# Stock Complexes {#sec-complexes}

```{r}
#| echo: false
#| include: false
library(openMSE)
```

In `openMSE`, every stock belongs to a stock complex. When each stock is managed independently, no complex configuration is needed. Multi-stock complexes are used when two or more stocks cannot be distinguished in the data: for example, males and females modelled as separate stocks, or two species landed together in a mixed fishery where catch cannot be attributed to either species individually. In these cases the stocks share a single combined data stream and receive a single joint management action such as a total allowable catch that applies to the complex as a whole.

The key consequences of grouping stocks into a complex are:

- **Combined `Obs`**: the `Obs` object is defined at the complex level, not the stock level. A single observation model applies to the aggregated data for the whole complex.
- **Combined data**: fishery observations (catch, indices, compositions) are aggregated across all stocks in the complex before being passed to the MP. The MP sees one `Data` object per complex.
- **Joint management advice**: the MP returns a single management action for the complex, which is then applied across the stocks within it.
- **Joint reference points**: $F_\text{MSY}$ is the single apical fishing mortality maximising total yield across all stocks in the complex (see @sec-refpoints).

Note that fleets remain stock-specific; each fleet has its own selectivity, retention, and effort schedule defined independently for each stock. Complexes operate above the fleet level, determining how data and advice are aggregated.

## The `Complexes` Slot

Complexes are defined via the `Complexes` slot of the `OM` object, which is a named list where each element is an integer vector of stock indices belonging to that complex:

```{r}
#| eval: false
# Two stocks in one complex
OM(
  Stock     = list(AlbacoreExStock, ButterfishExStock),
  Fleet     = AsympExFleet,
  Obs       = AgeStructuredObs,
  Complexes = list(StockComplex = 1:2)
)
```

Here both stocks (indices 1 and 2) are assigned to a single complex named `"StockComplex"`. The `Obs` object is indexed by complex and fleet, so a single observation model applies to the aggregated data for the whole complex.

When `Complexes` is not specified, `openMSE` sets a default automatically:

- If the `OM` contains a single stock, that stock forms its own complex.
- If the `OM` has a single `Data` object for multiple stocks, all stocks are placed in one complex.
- Otherwise, each stock is treated as its own independent complex.

```{r}
#| eval: false
# Inspect the complexes defined in an OM or Hist object
Complexes(myOM)
Complexes(Hist)
```

## Multi-Stock vs Multi-Complex

It is important to distinguish between a multi-stock, multi-complex OM and a multi-stock, single-complex OM:

- **Multi-stock, multi-complex**: each stock is managed independently, with its own `Obs`, its own `Data` object, and its own management advice. This is the default when `Complexes` is not specified.
- **Multi-stock, single complex**: all stocks share one `Obs`, their data are aggregated into one `Data` object, and management advice is applied jointly. This requires explicitly setting `Complexes`.

The `MultiStockOM` example has two stocks each in their own complex; `ComplexOM` has two stocks in a single shared complex.

```{r}
#| eval: false
# Two stocks, each managed separately (two complexes)
Complexes(MultiStockOM) # NULL

Complexes(Populate(MultiStockOM)) # list length 2

# Two stocks managed jointly (one complex)
Complexes(ComplexOM)
```

## Stock-Recruitment Linkages

By default each stock's recruitment is driven by its own spawning production. The `SPFrom` slot of `Stock@SRR` can designate a different stock as the source of spawning output used to determine recruitment. The most common case is a sex-structured model where female spawning biomass drives recruitment for both sexes. `SPFrom` operates at the OM level and can reference any stock regardless of which complex it belongs to.

## Reference Points

`CalcMSY()` finds the single apical $F$ that maximises total yield across all stocks in the complex. Stock-level reference points (BMSY, SBMSY, etc.) are then evaluated at this complex-level $F_\text{MSY}$, reflecting the reality that a single management action applies to all stocks simultaneously. See @sec-refpoints for details.
