# Reference Points {#sec-refpoints}

```{r}
#| warning: false
#| message: false
#| echo: false
#| include: false
library(openMSE)
```

Reference points are benchmark values that describe the state of a stock or fishery under specified conditions. They provide the yardstick against which current stock status is measured and against which management strategy performance is evaluated. In `openMSE`, reference points are calculated from the operating model after the historical dynamics have been run, and are stored in `Hist@Reference` and can be accessed with `Reference(Hist)`.

Two broad categories of reference points are used in openMSE:

- **Unfished reference points:** quantities under no fishing, including unfished biomass ($B_0$), unfished spawning biomass ($SB_0$), unfished spawning production ($SP_0$), and unfished recruitment ($R_0$). These describe the baseline state of the stock and are calculated as part of the unfished equilibrium (see @sec-tech-unfished).
- **MSY-based reference points:** quantities at or derived from maximum sustainable yield, including $F_\text{MSY}$, $B_\text{MSY}$, $SB_\text{MSY}$, $SP_\text{MSY}$, $F_{0.1}$, $F_\text{Max}$, $F_\text{crash}$, and associated SPR quantities.

Additional reference points include the mean generation time (MGT), the reference yield under a fixed-$F$ policy, and $B_\text{Low}$ (a rebuilding-based limit reference point used in DFO Canada's Precautionary Approach framework).

## MSY Reference Points

MSY reference points are calculated automatically by `Simulate()` and stored in `Hist@Reference@MSY`. They can also be recalculated at any time using `CalcMSY()`, which takes a `Hist` or `OM` object and returns a `refpointsMSY` object.

```{r}
#| eval: false
Hist <- Simulate(myOM)

# Full reference object
Reference(Hist)

# Full refpointsMSY object
MSYRefs(Hist)

# Individual quantities
FMSY(Hist)
BMSY(Hist)
SBMSY(Hist)
MSYLandings(Hist)
```

By default, reference points are evaluated using parameters from the last historical year. Alternative years can be specified via the `Years` argument:

```{r}
#| eval: false
# Reference points at a specific year
CalcMSY(Hist, Years = 2010)

# Reference points at multiple years (evaluated independently)
CalcMSY(Hist, Years = c(2005, 2010, 2020))
```

Evaluating reference points at multiple years is useful for examining how $F_\text{MSY}$ or $B_\text{MSY}$ shift as biological or fishery parameters change over time.

`CalcMSY()` returns a `refpointsMSY` object with the following slots, all with dimensions `Sim × Stock × Year`:

| Slot | Description |
|------|-------------|
| `FMSY` | Apical fishing mortality at MSY |
| `BMSY` | Total biomass at MSY |
| `SBMSY` | Spawning biomass at MSY |
| `SPMSY` | Spawning production at MSY |
| `SPRMSY` | Spawning potential ratio at MSY |
| `MSYLandings` | Landed catch at MSY |
| `MSYDiscards` | Dead discards at MSY |

### Removals vs Landings

`CalcMSY()` can maximise either total removals (landings plus dead discards) or landings alone, controlled by the `type` argument:

```{r}
#| eval: false
CalcMSY(Hist, type = "Removals")  # default
CalcMSY(Hist, type = "Landings")
```

Removals-based MSY finds the $F$ that maximises total fish removed from the population; landings-based MSY finds the $F$ that maximises retained catch. The two diverge when discards are substantial: the $F$ that maximises landings may be lower than the $F$ that maximises total removals, because at high $F$ a greater proportion of removals are discarded fish that contribute to mortality but not to yield.

### Multi-Stock Complexes

When a stock complex contains more than one stock, `CalcMSY()` finds the single $F_\text{MSY}$ maximising total yield across all stocks. Stock-level quantities (BMSY, SBMSY, etc.) are then evaluated at this complex-level $F_\text{MSY}$. See @sec-tech-refpoints for details.

## Yield-Per-Recruit Reference Points

::: {.callout-note}
$F_\text{Max}$ and $F_{0.1}$ are not yet calculated automatically. Code to compute and store these reference points is under development.
:::

Two $F$-based reference points are derived from the yield-per-recruit (YPR) curve and do not require a stock-recruitment relationship:

- **$F_\text{Max}$:** the apical fishing mortality that maximises YPR.
- **$F_{0.1}$:** the apical fishing mortality at which the slope of the YPR curve is 10% of its initial slope.

## Spawning Potential Ratio Reference Points

::: {.callout-note}
$F_\text{crash}$ and $\text{SPR}_\text{crash}$ are not yet calculated automatically. Code to compute and store these reference points is under development.
:::

The spawning potential ratio $\text{SPR}(F) = \Phi_F / \Phi_0$ measures the fraction of unfished spawning production per recruit remaining at fishing mortality $F$. Two SPR-based reference points are computed:

- **$F_\text{crash}$:** the fishing mortality at which equilibrium recruitment falls to zero.
- **$\text{SPR}_\text{crash}$:** the spawning potential ratio at $F_\text{crash}$, i.e. the SPR below which equilibrium recruitment is zero.

## Reference Yield

The reference yield provides a practical benchmark for comparing MP performance. It is the highest yield (landings or removals) achievable under a fixed-$F$ policy over the projection period, accounting for the actual simulated population trajectory and recruitment variability. It is computed separately for landings and removals, stored in `Hist@Reference@Landings` and `Hist@Reference@Removals`, and calculated automatically during `Simulate()` (landings by default; removals optionally).

For seasonal models, the effort pattern from the last complete historical year (all $S$ seasons) is extracted as the baseline. A single scalar multiplier is then optimised to scale all fleets and all seasons simultaneously, preserving the relative seasonal effort distribution while searching over the overall level of fishing pressure.

## Mean Generation Time

The mean generation time (MGT) is the mean age of a mature female in the population under natural mortality only at unfished equilibrium:

$$\text{MGT} = \frac{\sum_a a \cdot l_a \cdot m_a}{\sum_a l_a \cdot m_a}$$

where $l_a$ is survivorship to age $a$ under natural mortality only and $m_a$ is maturity-at-age. MGT is used as the time unit for the $B_\text{Low}$ calculation below and is computed by `CalcMGT()`:

```{r}
#| eval: false
CalcMGT(Hist)

# At a specific year
CalcMGT(Hist, Years = 2010)
```

MGT is stored in `Hist@Reference@MGT` and has dimensions `Sim × Stock × Year`. In seasonal models, age classes are in decimal-year units, so MGT is returned in years regardless of seasonal resolution.

## B~Low~

::: {.callout-note}
$B_\text{Low}$ is not calculated by default due to its computational cost. It must be enabled as an argument to `Simulate()`. Code for this is under development.
:::

$B_\text{Low}$ is a rebuilding-based limit reference point used in DFO Canada's Precautionary Approach framework. It is defined as the current spawning biomass below which a population, if it were to recover with zero fishing, could not rebuild to $B_\text{frac} \times SB_\text{MSY}$ within $\text{HZN} \times \text{MGT}$ years. The default values are $B_\text{frac} = 0.5$ and $\text{HZN} = 2$.

Conceptually, if current spawning biomass falls below $B_\text{Low}$, then even with immediate cessation of fishing the stock would require more than two mean generation times to recover to half of $SB_\text{MSY}$.

$B_\text{Low}$ is stored in `Hist@Reference@BLow`. Because it requires a forward simulation of the population trajectory, it is computationally expensive and is not calculated by default; it is enabled via an argument to `Simulate()`.

## Per-Recruit Curves

The underlying per-recruit calculations used to derive $F_\text{MSY}$, $F_{0.1}$, $F_\text{Max}$, and $F_\text{crash}$ can be examined directly using `CalcPerRecruit()` and `CalcEquilibrium()`. These are useful for understanding yield curves, plotting SPR against $F$, and diagnosing model behaviour.

```{r}
#| eval: false
PR <- CalcPerRecruit(Hist, apicalF = seq(0, 1, by = 0.05))
Eq <- CalcEquilibrium(Hist)
```

Both functions accept a `Years` argument and a `Complex` argument.

## Depletion and Stock Status

Current stock status is typically expressed as depletion, the ratio of current biomass to a reference level:

- $B/B_0$: current total biomass relative to unfished biomass
- $SB/SB_0$: current spawning biomass relative to unfished spawning biomass
- $SP/SP_0$: current spawning production relative to unfished spawning production
- $B/B_\text{MSY}$: current biomass relative to $B_\text{MSY}$
- $F/F_\text{MSY}$: current fishing mortality relative to $F_\text{MSY}$

Spawning biomass ($SB$) and spawning production ($SP$) measure related but distinct quantities. $SB$ is the biomass of reproductively active fish (maturity-at-age $\times$ weight-at-age $\times$ numbers). $SP$ is fecundity-weighted abundance, the quantity that directly drives the stock-recruitment relationship. When fecundity is proportional to weight, $SB$ and $SP$ are proportional and give the same depletion ratio; they diverge when fecundity is specified independently of weight (e.g. via age-specific egg production).

Helper functions for extracting these quantities from a `Hist` or `MSE` object:

```{r}
#| eval: false
# Relative to unfished
B_B0(Hist)       # total biomass / B0
SB_SB0(Hist)     # spawning biomass / SB0
SP_SP0(Hist)     # spawning production / SP0

# Relative to MSY
B_BMSY(Hist)     # total biomass / BMSY
SB_SBMSY(Hist)   # spawning biomass / SBMSY
SP_SPMSY(Hist)   # spawning production / SPMSY
F_FMSY(Hist)     # total apical F (summed over fleets) / FMSY
```

All functions accept a `df` argument (`TRUE` (default) returns a data frame; `FALSE` returns an array).

## Seasonal Models {#sec-concept-refpoints-seasonal}

Seasonal operating models subdivide each calendar year into $S$ seasons with season-specific biology, mortality, and recruitment. Reference points for seasonal models are expressed on an annual scale and follow the same logic as annual models. The `Years` argument to `CalcMSY()`, `CalcPerRecruit()`, and `CalcEquilibrium()` should specify complete calendar years (integers); all seasons within each requested year are used automatically.

### Annual reference points from seasonal dynamics

All reference point quantities are reported as annual values. $B_\text{MSY}$, $SB_\text{MSY}$, and $SP_\text{MSY}$ represent equilibrium population state at $F_\text{MSY}$; MSY is total annual catch across all seasons; SPR is the ratio of fished to unfished annual spawning production per recruit.

In a seasonal model, biomass fluctuates within the year as seasonal mortality and recruitment pulses occur. The per-recruit framework accounts for this by computing each quantity as a recruitment-weighted average over birth-season cohorts (see @sec-tech-refpoints-seasonal for the full equations).

### Cohorts and birth seasons

The seasonal recruitment pattern is controlled by `SRR@R0`: the proportion of annual recruitment entering in season $s_0$ is $\pi_{s_0} = R_{0,s_0} / R_{0,\text{annual}}$. Per-recruit quantities (spawning production, biomass, yield) are computed for each birth season separately and combined using these $\pi_{s_0}$ weights. Full equations are given in @sec-tech-refpoints-seasonal.

Cohorts born in different seasons experience different lifetime mortality and fecundity trajectories, because they encounter each calendar season at different ages. The $\pi_{s_0}$ weighting correctly averages over the full composition of annual recruitment across birth seasons.

### Fecundity and spawning production

`AdjustSeasonalFecundity()` (called automatically during `Simulate()`) scales fecundity-at-age by season in proportion to the seasonal recruitment pattern. This means $\Phi_0$ (unfished spawning production per recruit, the sum of fecundity-weighted survival over all ages and seasons) reflects the actual timing and magnitude of spawning across the year, consistent with whatever recruitment seasonality is specified. SPR $= \Phi_F / \Phi_0$ retains its standard interpretation: the fraction of unfished spawning production per recruit remaining at fishing mortality $F$, regardless of whether spawning is concentrated in one season or distributed across several.

### Interpretation of $F_\text{MSY}$

$F_\text{MSY}$ is the annual apical fishing mortality, the sum of instantaneous rates experienced by the most heavily fished age class across all seasons in a year. This is directly comparable to $F_\text{MSY}$ from an annual model. The optimisation scales a peak-season effort scalar, with all other seasons proportional to it via the historical effort pattern; $F_\text{MSY}$ is reported as the corresponding annual total.

## Spatial Models {#sec-concept-refpoints-spatial}

For spatially structured models, reference points are currently calculated using spatially-aggregated parameters and should be interpreted as spatial averages. A spatially-explicit equilibrium approach is planned as a future option.

For full technical details of all reference point calculations, see @sec-tech-refpoints.
