# Spatial Structure {#sec-spatial}

```{r}
#| echo: false
#| include: false
library(MSEtool)
rnorm(5)
```

`openMSE` supports spatially structured operating models with an arbitrary number of areas. Fish move between areas according to a movement probability matrix, and fishing effort can be distributed across areas either dynamically or according to a fixed allocation (see @sec-spatial-effort). Management can be applied differentially by area through spatial closures, and area-specific selectivity and retention.
 
The `Spatial` subcomponent of a `Stock` object defines the spatial structure. It is optional; if omitted the model assumes a single well-mixed area with no spatial dynamics.

This chapter describes the available approaches for specifying spatial structure, from the simplest two-area case through to models with age-varying movement and three or more areas. All `Stock` objects within an operating model must share the same spatial configuration (i.e., the same number of areas).

## Single-Area Models

For a non-spatial model, simply omit the `Spatial` subcomponent entirely. `Populate()` will create a default single-area structure automatically:

```{r}
# No Spatial subcomponent: single well-mixed area assumed
stock <- Stock(Name = "My Stock")
# Spatial(stock) is NULL; Populate() will assign a single-area default
```

## Two-Area Models

Two-area models are a common spatial configuration. They require three inputs: the unfished biomass distribution across areas, the probability of remaining in an area, and the relative size of each area.

For two-area models, `UnfishedDist` and `RelativeSize` each specify Area 1 only; the corresponding Area 2 values are derived as `1 - x`. Similarly, `ProbStaying` specifies Area 1 only; the Area 2 staying probability is solved internally so that the asymptotic movement equilibrium reproduces the target `UnfishedDist`.

Each parameter accepts a scalar (constant across all simulations), a length-2 bounds vector (sampled from `Uniform(lower, upper)` independently for each simulation), or a full `Sim × Area × Age × Year` array:

```{r}
# Stochastic samples from uniform distribution
Spatial(stock) <- Spatial(
  UnfishedDist = c(0.2, 0.4),   # fraction of unfished biomass in Area 1
  ProbStaying  = c(0.6, 0.8),   # probability of staying in Area 1
  RelativeSize = c(0.3, 0.5)    # relative habitat size of Area 1
)
```

```{r}
# Fixed — same across all simulations
Spatial(stock) <- Spatial(
  UnfishedDist = 0.3,
  ProbStaying  = 0.7,
  RelativeSize = 0.4
)
```

## Relative Area Size

`RelativeSize` controls how biomass is converted to density for the purpose of spatial effort allocation. Fleets are attracted to areas with high biomass density (biomass per unit area), so `RelativeSize` affects where effort concentrates even when biomass is held constant.

Three options are available:

### Fixed proportion

For two-area models, a scalar or bounds vector gives the fraction of total habitat in Area 1, with Area 2 taken as `1 - RelativeSize`:

```{r}
# Area 1 is 40% of total habitat
RelativeSize = 0.4   
```

For models with three or more areas, this shortcut doesn't apply; `RelativeSize` must instead be supplied as a vector of length `nArea` giving every area's share directly, summing to 1:

```{r}
# Areas 1-3 are 10%, 40%, 50% of total habitat
RelativeSize = c(0.1, 0.4, 0.5)   
```

### Equal density

Setting `RelativeSize = "EqualDensity"` sets each area's relative size equal to its mean unfished biomass share across ages and years. This ensures that unfished biomass density is equal across areas, regardless of the number of areas:

```{r}
Spatial(stock) <- Spatial(
  UnfishedDist = 0.3,
  ProbStaying  = 0.7,
  RelativeSize = "EqualDensity"
)
```

### Equal size

When `RelativeSize` is omitted (`NULL`), all areas are assumed equal in size (`1 / nArea` each), whatever the number of areas.

## How the Movement Matrix is Derived

When `UnfishedDist` and `ProbStaying` are supplied, `Populate()` derives the movement probability matrix by numerical optimisation. For each simulation, age, and year, the matrix diagonal (staying probabilities) and off-diagonal elements (movement probabilities) are fitted to simultaneously satisfy:

1. The asymptotic distribution implied by the movement matrix matches the target `UnfishedDist`.
2. The diagonal elements match the target `ProbStaying`.

For two-area models this system has a closed-form solution. For models with three or more areas, a penalty-based optimisation is used (see @sec-spatial-multiarea). The full mathematical specification of both cases is given in @sec-tech-movement-derivation.

The resulting `Movement` array has dimensions `Sim × FromArea × ToArea × Age × Year`. Row `[i, ]` gives the probabilities of moving from area `i` to each other area, and rows sum to 1. See @sec-tech-historical-number for how the resulting movement matrix is applied within the historical population dynamics.

## Age-Varying Movement

Movement rates can vary by age by supplying `UnfishedDist` and/or `ProbStaying` as arrays with an `Age` dimension. A separate movement matrix is fitted for each age class: the matrix for age `a` is fit so that applying it to age `a`'s target unfished distribution reproduces age `a + 1`'s target distribution, chaining the age-specific matrices so the full age-specific distribution profile is realised as fish age through the model. The plus group, and any model where `UnfishedDist` doesn't vary by age, fits to the matrix's own asymptotic distribution instead, since the same matrix is reapplied every time step.

An example application is ontogenetic habitat shift, where juveniles and adults occupy different areas:

```r
ages <- Ages(MaxAge = 10)
nage <- nAge(ages)

# Juveniles concentrate in Area 1; adults spread to Area 2
area1_frac <- seq(0.8, 0.2, length.out = nage)

ud <- array(
  c(area1_frac, 1 - area1_frac),
  dim      = c(1, 2, nage),
  dimnames = list(Sim = 1, Area = 1:2, Age = Classes(ages))
)

Spatial(stock) <- Spatial(
  UnfishedDist = ud,
  ProbStaying  = 0.9
)
```

## Multi-Area Models {#sec-spatial-multiarea}

Models with three or more areas require `UnfishedDist` to be supplied as an array summing to 1 across the `Area` dimension, and additionally require `FracOther` to guide the movement optimisation.

`FracOther` is a `Sim × FromArea × ToArea` array (optionally extended with `Age` and `Year` dimensions) where off-diagonal element `[i, j]` gives the relative probability of moving from area `i` to area `j`. Diagonal elements must be `NA` as the staying probability is taken from `ProbStaying`. The values in `FracOther` should sum to 1, but are normalised internally so it is not essential.

```r
nArea <- 3

# Unfished biomass: 50% in Area 1, 20% in Area 2, 30% in Area 3
ud <- matrix(c(0.5, 0.2, 0.3), nrow = 1, ncol = nArea)

# FracOther: relative movement probabilities between areas
# Area 1 mostly connects to Area 2, weakly to Area 3
# Area 2 connects equally to Areas 1 and 3
# Area 3 mostly connects to Area 2, weakly to Area 1
fo <- array(NA, dim = c(1, nArea, nArea))
fo[1, 1, ] <- c(NA,  1,   0.1)
fo[1, 2, ] <- c(1,   NA,  1  )
fo[1, 3, ] <- c(0.1, 1,   NA )

Spatial(stock) <- Spatial(
  UnfishedDist = ud,
  ProbStaying  = c(0.9, 0.2, 0.9),
  FracOther    = fo,
  RelativeSize = c(0.1, 0.4, 0.5)
)
```

See @sec-tech-movement-derivation for the technical details regarding how the movement matrix is calculated.

## Supplying a Movement Matrix Directly

As an alternative to specifying `UnfishedDist` and `ProbStaying` and letting `Populate()` derive the movement matrix, the movement matrix can be supplied directly. This is useful when the matrix has been estimated externally (e.g., from a tagging study or a spatially-explicit model).

The matrix must have dimensions `Sim × FromArea × ToArea`, optionally extended to `Sim × FromArea × ToArea × Age × Year`. When `Movement` is supplied, the asymptotic unfished distribution is back-calculated from the matrix and assigned to `UnfishedDist`, overwriting any values already present:

```r
mov <- array(
  c(0.7, 0.3,
    0.2, 0.8),
  dim      = c(1, 2, 2),
  dimnames = list(Sim = 1, FromArea = 1:2, ToArea = 1:2)
)

Spatial(stock) <- Spatial(Movement = mov)
```

## Spatial Effort Allocation {#sec-spatial-effort}

Spatial structure in the `Stock` object defines where fish are. How fishing effort is distributed across areas is controlled by the `Effort` subcomponent of each `Fleet` object, specifically the `Distribution` and `Targeting` slots.

When `Distribution` is not specified, effort is allocated dynamically each year based on relative exploitable biomass density across areas, with the degree of concentration controlled by the `Targeting` parameter. A value of 0 produces uniform effort allocation regardless of biomass distribution; larger values concentrate effort more strongly in the highest-density area. The default `Targeting` value is 0.8.

Users can fix effort to a specific spatial allocation by supplying `Distribution` as a `Sim × Year × Area` array, bypassing the dynamic allocation algorithm entirely. Any cells set to a non-`NA` value are treated as fixed overrides and are not modified.

See `?Effort` for full documentation of `Distribution` and `Targeting`, and @sec-tech-historical-effort for the mathematical specification of the spatial effort allocation algorithm. 
