# Populating an Operating Model {#sec-populate}

```{r}
#| echo: false
#| include: false
library(MSEtool)
```

Building an operating model in `openMSE` is a two-stage process. First, the user constructs the `OM` object and its sub-components by specifying parameter values as described in @sec-primaryobjects and @sec-schedules. Second, the `Populate()` function expands these compact inputs into the full arrays (e.g., `Sim × Age × Year`), samples any stochastic values, resolves model functions, builds derived quantities, and and checks the `OM` for internal consistency.

In normal use, `Populate()` is called automatically when running `Simulate()`, so most users will not need to call it directly. It can be useful, however, for inspecting what a set of parameter values produces before assembling a full operating model.

## `Populate()`

`Populate()` is a generic dispatcher. It inspects the class of the object passed to it and calls the appropriate class-specific population function:

```{r}
#| eval: false

# Populate a full operating model
OM <- Populate(OM)

# Populate a stock object independently
Stock <- Populate(Stock, nSim = 100, nYear = 40, pYear = 30)

# Populate a fleet using an already-populated stock
Fleet <- Populate(Fleet, Stock = Stock)
```

The primary class-specific functions called by `Populate()` are: 

- `PopulateOM()`
- `PopulateStock()`
- `PopulateFleet()`
- `PopulateObs()`
- `PopulateImp()`

These functions in turn call the population functions for their individual subcomponents (`PopulateLength()`, `PopulateSelectivity()`, etc.). See `?PopulateOM` and related help pages for more information.

## Using `Populate`

Although `Populate()` doesn't need to be called directly, individual objects can be populated independently. This is sometimes useful for inspecting what a set of parameter values produces before assembling a full operating model, for example inspecting the length-at-age schedule produced by a set of parameters:

```{r}
stock <- Populate(AlbacoreExStock, nSim = 100, nYear = 30, pYear = 20)
Length(stock) |> MeanAtAge() |> head()
```

When populating a `Fleet` object, the paired `Stock` object must be provided so that quantities such as the length structure (e.g., the age-length key required for size-to-age conversions) are available: 

```{r}
#| eval: false
fleet <- Populate(AsympExFleet, Stock = stock)
```

## Seeding and Reproducibility

Each subcomponent population function receives its own integer seed, derived by incrementing the base seed stored in `OM@Seed`. This ensures that:

- All stochastic draws are fully reproducible given a fixed `OM@Seed`.
- Population of one subcomponent does not affect the random state of another.
- All sub-objects (e.g., `Stock`, `Fleet`, etc) receive distinct seeds, so stochastic variation in one component is independent of another even when the same parameter ranges are specified.

To change the stochastic realisation of an operating model, update `OM@Seed` and call `Populate()` again:

```r
Seed(OM) <- 42
OM <- Populate(OM)
```

