# An Example MSE

```{r}
#| warning: false
#| message: false
#| echo: false
#| include: false
library(openMSE)
rnorm(5)
```

This chapter provides a walkthrough of the MSE process in `openMSE`, from specifying an operating model through to evaluating management procedure performance. The aim is for users to see the process end-to-end and become familiar with how the primary objects and functions interact.

## Specify an Operating Model

The first step in any MSE is to construct the operating model. The OM is arguably the most important component of the analysis, it represents the best available understanding (or one hypothesis of several alternative understandings) of the stock and fishery dynamics, and all management procedure evaluations are conditioned on it.

This chapter uses `SingleStockOM`, one of `openMSE`'s built-in example operating models. It is a single-stock, single-fleet OM with `r nSim(SingleStockOM)` stochastic simulations, `r nYear(SingleStockOM)` historical years, and `r pYear(SingleStockOM)` projection years. 

The small number of simulations keeps the model fast to run for demonstration purposes. In practice the number of simulations should be sufficiently high that the quantitative performance metrics are stable. 

See @sec-exampleobjects for an overview of all example objects and [Developing Operating Models](developing-oms.qmd) for details on the different methods for constructing operating models.

Typing the object name in the console prints a summary of its contents (@fig-output-singlestockom):

```{r}
#| eval: false
SingleStockOM
```

![Summary of the `SingleStockOM` object printed to the R console.](images/output-singlestockom){#fig-output-singlestockom width=80%}

More information on `SingleStockOM` can be found by accessing the help documentation:
```{r}
#| eval: false
?SingleStockOM
```

and more information on `OM` objects can be found in @sec-overview as well as the help documentation:
```{r}
#| eval: false
?OM
```

## Simulate Historical Fishery

```{r}
#| include: false
knitr::read_chunk("R/example-mse.R")
```

Once an OM has been constructed, the historical fishery can be simulated (or reconstructed) using the `Simulate` function:

```{r run-hist}
#| echo: false
```
```{r}
#| eval: false
Hist <- Simulate(SingleStockOM)
```

As with the `OM` object above, typing the object name in the console prints a summary of its contents (@fig-output-singlestock-hist):

![Summary of the `Hist` object printed to the R console.](images/output-singlestock-hist){#fig-output-singlestock-hist width=80%}

The `Hist` object contains the complete historical simulation output. Its key components are:

- **The OM**: the `OM` object used to generate the simulation (`OM`).
- **Reference quantities**: unfished equilibrium and dynamic trajectories (`Unfished`) and biological reference points (`Reference`). See @sec-refpoints for more details.

- **Population dynamics**: numbers-at-age (`Number`), total biomass (`Biomass`), spawning biomass (`SBiomass`), and spawning production (`SProduction`), all indexed by simulation, stock, and year. Because stocks can have different numbers of age classes, any output with an age or size dimension is stored as a nested list by stock (or by stock and fleet), rather than a single array.
- **Fishery dynamics**: landings and discards (total, at-age, and at-size), fishing effort and its spatial distribution, and fishing mortality decomposed into interacting, dead, and retained components, all indexed by simulation, stock, fleet, year, and area.
- **Observed data**: the historical fishery data for each stock complex (`Data` object; not shown in figure above). Either the real data provided in `OM@Data` and/or simulated data generated by the observation model during the historical period. See @sec-obsdataconditioning for more details.

See `?Hist` for full documentation of the `hist-class` and its slots.

### Exploring the `Hist` Object

The historical fishery dynamics contained in `Hist` can be explored using a set of extractor functions that return tidy data frames. The key functions are:

- `Number()`: total abundance (numbers) across all ages and areas.
- `Biomass()`: total biomass. Optionally by age and/or area.
- `SBiomass()`: spawning biomass. Optionally by age and/or area.
- `SProduction()`: spawning production (egg production if `Fecundity` is specified, otherwise equal to spawning biomass; summed over ages and areas).
- `Landings()`: retained catch in biomass (or numbers if `byAge = TRUE`). Optionally by age, size, area, and/or fleet.
- `Discards()`: dead discarded catch. Same structure as `Landings`.
- `Removals()`: total removals (landings plus dead discards). Same structure as `Landings`.
- `FInteract()`: apical fishing mortality for all fish that encounter the gear, including those that are released alive. Optionally by age, size, area, and/or fleet.
- `FDead()`: apical fishing mortality for fish killed by the gear (retained plus dead discards). Same structure as `FInteract`.
- `FRetain()`: apical fishing mortality for retained fish only. Same structure as `FInteract`.
- `Effort()`: fishing effort by fleet.

Each accepts a `Hist` object and returns a data frame with columns `Sim`, `Stock`, `Year`, `Period`, and `Value`, with optional arguments to retain the `Fleet`, `Age`, `Size`, and `Area` dimensions.

See `?Biomass`, `?Landings`, `?Number`, and `?FDead` for full documentation.

```{r explore-hist-biomass}
```

Landings and discards are extracted similarly:

```{r explore-hist-catch}
```


::: {.callout-tip}
## Array Dimensions 
By default, arrays stored in `Hist` are automatically compressed along the `Sim` dimension to the minimum number of unique simulations (typically 1 or `nSim`), rather than storing a full `nSim` copy for every replicate. When many simulations share identical parameter draws, as can happen when stochastic variation is low or parameters are fixed, this compression can substantially reduce memory use with no loss of information.

However, some operations, such as plotting one line per simulation, may require the `Sim` dimension to be exactly `nSim`. In these cases, use `Extend()` to expand an array back to the full simulation dimension:

```{r array-dimensions}

```

See `?Extend` and `?ReduceDims` for full documentation.
:::

## Specify Management Procedures

Once the historical fishery dynamics have been simulated, the operating model can be projected forward under different management policies, known as management procedures (MPs).

The design and selection of MPs is a central component of any MSE, and can range from simple empirical rules based on catch history to complex model-based procedures involving stock assessment. This example MSE uses a set of simple example MPs included with `openMSE` (see @sec-exampleMPs):

- `CurrentEffort`: maintains fishing effort at its terminal historical level for all fleets.
- `CurrentCatch:` sets the TAC for each fleet to the total removals (landings plus discards) observed in the last historical year.
- `AverageCatch`: sets the TAC to the mean total removals across all historical years.

A full description of MP development and specification is given in @sec-managementprocedures.

## Run Forward Projections 

Once the historical dynamics have been simulated and the MPs specified, the closed-loop projections are run using `Project()`:

```{r forward-projections}
#| cache: true
#| echo: false
```

```{r forward-projections2}
#| eval: false
#| 
```

`Project()` runs each MP in turn over the projection period. At each management interval, the operating model passes the current `Data` object to the MP, applies the returned `Advice` to the simulated population, and advances the model forward. This loop is repeated until the end of the projection period.

As with the `OM` and `Hist` objects, typing the object name in the console prints a summary of its contents (@fig-output-singlestock-mse):

![Summary of the `MSE` object printed to the R console.](images/output-singlestock-mse){#fig-output-singlestock-mse width=80%}

The `MSE` object contains the complete projection output. It extends the `Hist` object with projection-period equivalents of all the population and fishery dynamics arrays, indexed by MP in addition to simulation, stock, fleet, year, and area. 

The same extractor functions used to explore `Hist` (`Biomass()`, `Landings()`, `FDead()`, etc.) work on `MSE` objects, with the historical and projection periods labelled via a `Period` column, and an additional `MP` column.

See `?MSE` for full documentation of the `mse-class` and its slots.

### Exploring the `MSE` Object

The projection results can be explored using the same extractor functions as for `Hist`:

```{r explore-mse-biomass}
```

### Evaluating Performance

MSE results are typically evaluated using quantitative performance metrics (PMs), numerical summaries of management procedure performance against objectives specified by fishery managers. 

Common examples include the probability of maintaining spawning biomass above a limit reference point, the average catch over the projection period, and the inter-annual variability in catch. PMs translate the raw simulation output into quantitative quantities that can be compared across MPs and used by managers to identify the MP that is most likely to meet their management objectives.

`openMSE` does not currently include built-in example PMs, but these are planned for a future release.

For interactive visualization and exploration of MSE results, we recommend [Slick](https://slick.bluematterscience.com/), a web-based application developed for summarizing and communicating MSE output to scientists, managers, and other stakeholders. 

Slick aprovides an interactive interface for plotting performance metrics, comparing MPs across objectives, and producing presentation-ready figures. It is particularly useful for stakeholder engagement and for the performance trade-off analysis that is central to the MP selection process.

