# An Example MSE {#sec-example-mse}

```{r}
#| warning: false
#| message: false
#| echo: false
#| include: false
library(openMSE)
rnorm(5)
```

This chapter provides a walkthrough of the MSE process, from specifying an operating model through to evaluating management procedure performance. The aim is for users to see the process end-to-end and become familiar with how the primary objects and functions interact.

## Specify an Operating Model

The first step in any MSE is to construct the operating model. The OM represents the best available understanding (or one hypothesis of several alternative understandings) of the stock and fishery dynamics, and all management procedure evaluations are conditioned on it.

This chapter uses `SingleStockOM`, one of `openMSE`'s built-in example operating models. It is a single-stock, single-fleet OM with `r nSim(SingleStockOM)` stochastic simulations, `r nYear(SingleStockOM)` historical years, and `r pYear(SingleStockOM)` projection years. 

The small number of simulations keeps the model fast to run for demonstration purposes. In practice the number of simulations should be sufficiently high that the quantitative performance metrics are stable. 

See @sec-exampleobjects for an overview of all example objects and [Developing Operating Models](developing-oms.qmd) for details on the different methods for constructing operating models.

Typing the OM object name in the console prints a summary of its contents (@fig-output-singlestockom):

```{r}
#| eval: false
SingleStockOM
```

![Summary of the `SingleStockOM` object printed to the R console.](images/output-singlestockom){#fig-output-singlestockom width=80%}

More information on `SingleStockOM` can be found by accessing the help documentation:
```{r}
#| eval: false
?SingleStockOM
```

and more information on `OM` objects can be found in @sec-overview as well as the help documentation:
```{r}
#| eval: false
?OM
```

## Simulate Historical Fishery

```{r}
#| include: false
knitr::read_chunk("R/example-mse.R")
```

Once an OM has been constructed, the historical fishery can be simulated (or reconstructed) using the `Simulate` function:

```{r run-hist}
#| echo: false
```
```{r}
#| eval: false
myHist <- Simulate(SingleStockOM)
```

As with the `OM` object above, typing the object name in the console prints a summary of its contents (@fig-output-singlestock-hist):

```{r}
#| eval: false
myHist
```

![Summary of the `Hist` object printed to the R console.](images/output-singlestock-hist){#fig-output-singlestock-hist width=80%}

The `Hist` object contains the complete historical simulation output. Its key components are:

- **The OM**: the `OM` object used to generate the simulation.
- **Reference quantities**: unfished equilibrium and dynamic trajectories (`Unfished`) and biological reference points (`Reference`). See @sec-refpoints for more details.
- **Population dynamics**: total biomass (`Biomass`), spawning biomass (`SBiomass`), and spawning production (`SProduction`), each `Sim × Stock × Year`. Numbers-at-age (`Number`) is a list of `Sim × Age × Year × Area` arrays, one per stock.
- **Fishery dynamics**: fishing effort and its spatial distribution (`Effort`, `Distribution`); total biomass interacting with the gear (`Interactions`); landings and discards, both totals and at-age/at-size (`Landings`, `Discards`, `LandingsAtAge`, `DiscardsAtAge`, `LandingsAtSize`, `DiscardsAtSize`); and fishing mortality decomposed into interacting, dead, and retained components (`FInteract`, `FDead`, `FRetain`), with area- and age-specific counterparts (`FInteractArea`, `FDeadArea`, `FRetainArea`).
- **Observed data**: the historical fishery data for each stock complex (`Data`; not shown in figure above). Either the real data provided in `OM@Data` and/or simulated data generated by the observation model during the historical period. See @sec-obsdataconditioning for more details.

::: {.callout-tip}
## Nested-List Outputs
Stocks can have different numbers of age classes, so any `Hist` output with an age dimension (e.g. `Number`, `LandingsAtAge`) is stored as a list by stock, rather than a single array, with `Fleet` retained as an array dimension where relevant.

At-size outputs (`LandingsAtSize`, `DiscardsAtSize`) are nested one level further, by stock *then* fleet, since fleets are not required to share a common size-class grid.
:::

See `?Hist` for full documentation of the `hist-class` and its slots.

### Exploring the `Hist` Object

The historical fishery dynamics contained in `Hist` can be explored using a set of extractor functions that return tidy data frames:

- `Number()`, `Biomass()`, `SBiomass()`, `SProduction()`: population abundance, total biomass, spawning biomass, and spawning production (egg production if `Fecundity` is specified, otherwise equal to spawning biomass).
- `Interactions()`, `Landings()`, `Discards()`, `Removals()`: total biomass interacting with the gear, retained catch, dead discards, and removals (landings plus dead discards).
- `FInteract()`, `FDead()`, `FRetain()`: apical fishing mortality for fish that interact with the gear, are killed by it, and are retained, respectively.
- `Effort()`: fishing effort by fleet.

Each returns a data frame with columns `Sim`, `Stock`, `Year`, `Period`, and `Value`, summed by default over age, size, area, and fleet. Optional `byAge`/`byArea`/`byFleet` arguments (and `bySize` for the catch-based functions) retain those dimensions instead.

See `?Biomass`, `?Landings`, `?Number`, and `?FDead` for full documentation. 

```{r explore-hist-biomass}
```

::: {.callout-tip}
## The `Units` Column
The `Units` column reported by `Biomass()`, `SBiomass()`, and `SProduction()` is computed by combining the stock's base unit (`Weight@Units` or `Fecundity@Units`) with the `SRR@Units` scaling factor applied to `R0` (see @sec-overview). `AlbacoreExStock` has `Weight@Units = "kg"` and `SRR@Units = 1000` (`R0` in thousands of fish), which resolves to `"t"`, since the two scaling factors compound.
:::

Landings and discards are extracted similarly:

```{r explore-hist-catch}
```

::: {.callout-tip} 
## Array Dimensions
By default, arrays stored in `Hist` are automatically compressed along the `Sim` dimension when the simulations all share identical values. For example, if all simulations produce zero discards, the discard array is stored with a single row rather than `nSim` identical rows.

Some operations, such as plotting one line per simulation, require the full `nSim` dimension. Use `Extend()` to expand an array back to full size:

```{r array-dimensions}
```

See `?Extend` and `?ReduceDims` for full documentation.
:::

## Specify Management Procedures 

Once the historical fishery dynamics have been simulated, the operating model can be projected forward under different management policies, known as management procedures (MPs).

The design and selection of MPs is a central component of any MSE, and can range from simple empirical rules based on catch history to complex model-based procedures involving stock assessment. 

This example MSE uses a set of simple example MPs included with `openMSE` (see @sec-exampleMPs), returned by the convenience function `ExampleMPs()`:

- `CurrentEffort`: maintains fishing effort at its terminal historical level for all fleets.
- `CurrentCatch`: sets the TAC for each fleet to the total removals observed in the last historical year.
- `AverageCatch`: sets the TAC to the mean total removals over the historical years.

A full description of MP development and specification is given in @sec-managementprocedures.

## Run Forward Projections 

Once the historical dynamics have been simulated and the MPs specified, the closed-loop projections are run using `Project()`:

```{r forward-projections}
#| cache: true
#| echo: false
```

```{r forward-projections2}
#| eval: false
```

`Project()` runs each MP in turn over the projection period. At each management interval, the operating model passes the current `Data` object to the MP, applies the returned `Advice` to the simulated population, and advances the model forward. This loop is repeated until the end of the projection period.

As with the `OM` and `Hist` objects, typing the object name in the console prints a summary of its contents (@fig-output-singlestock-mse):

![Summary of the `MSE` object printed to the R console.](images/output-singlestock-mse){#fig-output-singlestock-mse width=80%}

The `MSE` object contains the complete projection output. It extends the `Hist` object with projection-period equivalents of all the population and fishery dynamics arrays, indexed by MP in addition to simulation, stock, fleet, year, and area. 

The same extractor functions used to explore `Hist` (`Biomass()`, `Landings()`, `FDead()`, etc.) work on `MSE` objects, with the historical and projection periods labelled via a `Period` column, and an additional `MP` column.

See `?MSE` for full documentation of the `mse-class` and its slots.

### Exploring the `MSE` Object

The projection results can be explored using the same extractor functions as for `Hist`:

```{r explore-mse-biomass}
```

## Evaluating Performance

MSE results are typically evaluated using quantitative performance metrics (PMs), numerical summaries of management procedure performance against objectives specified by fishery managers. Common examples include the probability of maintaining spawning biomass above a limit reference point, the average catch over the projection period, and the inter-annual variability in catch. 

`openMSE` includes extractor functions for common derived quantities used in performance metric calculations. For example, `SB_SB0()` returns spawning biomass relative to unfished spawning biomass ($SB/SB_0$) for each simulation, MP, stock, and year (these extractor functions also work for `Hist` objects): 

A simple summary table across MPs:

```{r explore-mse-results}
```

### Plotting MSE Output

Projection results can also be visualised directly. `openMSE` provides a family of `Plot*()` functions (`PlotBiomass()`, `PlotSBiomass()`, `PlotF()`, etc.) that work on both `Hist` and `MSE` objects, so custom plotting code is rarely needed. @fig-mse-biomass uses `PlotSBiomass(myMSE, relative = "B0")` to show $SB/SB_0$ trajectories across simulations and MPs. See `?plot_hist` for the full family of plotting functions, including options to facet by stock/fleet, overlay individual simulation replicates, and plot relative to `BMSY` instead of `B0`.


```{r fig-mse-biomass}
#| echo: false
#| fig-cap: "Spawning biomass relative to unfished ($SB/SB_0$) across simulations
#|   and MPs for `SingleStockOM`, produced by `PlotSBiomass()`. The shaded region
#|   spans the 10th–90th percentile; the solid line shows the median. Historical
#|   years are shown in grey; each MP's projection is coloured and continues
#|   from the last historical value."
```



### Performance Metric Functions 

`openMSE` also includes a set of built-in performance metric functions (`PM_*`), which summarise MSE results against a specific management objective and return a `pm` object rather than a raw data frame. Each `pm` object reports the underlying statistic by simulation, stock, and MP (`@Stat`), and, for metrics with a pass/fail objective, the proportion of years that objective is met, averaged across simulations (`@Prob` and `@Mean`).

`PM_SBSBMSY()` and `PM_FFMSY()` give the probability of maintaining spawning biomass above SBMSY and fishing mortality below FMSY, respectively, over the projection period:

```{r pm-examples}
```

`PM_Yield()` reports mean projected yield instead. Yield has no pass/fail objective by default, so `@Prob` and `@Mean` are left `NA`; the projected yield itself is only in `@Stat`, which retains the `Sim` dimension and must be summarised across simulations directly:

```{r pm-yield}
```

See `?PM` for the full list of available performance metrics, and `?pm-class` for the structure of the returned object.

For interactive visualization and exploration of MSE results, we recommend [Slick](https://slick.bluematterscience.com/), a web-based application developed for summarizing and communicating MSE output to scientists, managers, and other stakeholders. Slick aprovides an interactive interface for plotting performance metrics, comparing MPs across objectives, and producing presentation-ready figures. It is particularly useful for stakeholder engagement and for the performance trade-off analysis that is central to the MP selection process.

