# Introduction

`openMSE` is a framework for building operating models, analyzing fishery data, and conducting Management Strategy Evaluation (MSE) for a wide range of fisheries.

This manual is intended for fisheries scientists and analysts who wish to use `openMSE` for all stages of an MSE, including:

- building and conditioning operating models;
- simulating historical fishery dynamics;
- developing and testing management procedures;
- conducting closed-loop simulation testing;
- summarizing and presenting results.

## About `openMSE`

`openMSE` is an umbrella R package comprising three core packages:

1. [MSEtool](https://MSEtool.openMSE.com) — the Management Strategy Evaluation toolkit. The core package in the collection, containing the primary functions for constructing operating models, simulating fishery dynamics, and running closed-loop MSE.

2. [DLMtool](https://DLMtool.openMSE.com) — the Data-Limited Methods toolkit. A library of data-limited management procedures for use with `MSEtool`.

3. [SAMtool](https://SAMtool.openMSE.com) — the Stock Assessment Methods toolkit. A library of model-based assessment methods and management procedures for data-intensive applications. Includes the Rapid Conditioning Model.

Loading `openMSE` into an R session automatically loads all three packages. Each package has its own documentation page (linked above); this manual focuses on the overall framework rather than individual functions.

`openMSE` was developed by Tom Carruthers, Quang Huynh, and Adrian Hordyk at [Blue Matter Science](https://www.bluematterscience.com/).

## Installing `openMSE`

We recommend installing `openMSE` using [`pak`](https://pak.r-lib.org/reference/pak.html), which automatically checks that all dependency packages meet the minimum version requirements:

```{r}
#| eval: false
# Install pak if required
if (!requireNamespace("pak", quietly = TRUE))
  install.packages("pak")

# Install openMSE and dependencies
pak::pak("openMSE")
```

To upgrade all packages to the latest versions:

```{r}
#| eval: false
pak::pak("openMSE", upgrade = TRUE)
```

Then load `openMSE` into your R session:

```{r}
#| eval: false
library(openMSE)
```

::: callout-note
## Minimum Version Requirements

The features described in this manual require the following package versions:

- `openMSE >= 2.0`
- `MSEtool >= 4.0`
- `DLMtool >= 7.0`
- `SAMtool >= 2.0`
:::

::: callout-important
## Development Version

`MSEtool v4.0` has not yet been released on CRAN and must be installed from GitHub:

```{r}
#| eval: false
pak::pkg_install("blue-matter/MSEtool")
```

All code in this manual works with this development version. The existing CRAN versions of `openMSE`, `DLMtool`, and `SAMtool` are compatible. This note will be removed once `MSEtool v4.0` is released on CRAN.
:::

## About Version 2.0

`openMSE 2.0` is a major update to the framework. Key changes include a complete refactoring of the source code for improved efficiency and maintainability, and a redesign of the core S4 object classes and their associated functions. The majority of these changes are in `MSEtool`.

`openMSE 2.0` is backwards compatible; code written for earlier versions should continue to work without modification. Please file an [issue](https://github.com/Blue-Matter/openMSE/issues) if you encounter any compatibility problems. Legacy code will receive fixes to maintain compatibility but will not otherwise be updated; for new work, we recommend following the methods described in this manual.

For a full description of changes see @sec-changes.

