# Introduction {#sec-introduction}

`openMSE` is a framework for building operating models, analyzing fishery data, and conducting Management Strategy Evaluation (MSE).

This manual is intended for fisheries scientists and analysts who wish to use `openMSE` for all stages of an MSE, including:

- building and conditioning operating models;
- simulating historical fishery dynamics;
- developing and testing management procedures;
- conducting closed-loop simulation testing;
- summarizing and presenting results.

## About `openMSE`

`openMSE` is an umbrella R package comprising three core packages:

1. [MSEtool](https://MSEtool.openMSE.com): the Management Strategy Evaluation toolkit. The core package in the collection, containing the primary functions for constructing operating models, simulating fishery dynamics, and running closed-loop MSE.

2. [DLMtool](https://DLMtool.openMSE.com): the Data-Limited Methods toolkit. A library of data-limited management procedures for use with `MSEtool`.

3. [SAMtool](https://SAMtool.openMSE.com): the Stock Assessment Methods toolkit. A library of model-based assessment methods and management procedures for data-intensive applications. Includes the Rapid Conditioning Model.

Loading `openMSE` into an R session automatically loads all three packages. Each package has its own documentation page (linked above); this manual focuses on the overall framework.

`openMSE` was developed by Adrian Hordyk, Quang Huynh, and Tom Carruthers at [Blue Matter Science](https://www.bluematterscience.com/).

## Prerequisites

To use `openMSE` or any of its component packages you must have:

1. A recent version of the R software installed on your machine;
2. While not essential, we highly recommend using an integrated development environment (IDE) such as RStudio;
3. The latest version of `openMSE` and the individual component packages installed.

### Installing R
The R statistical language and environment can be freely downloaded from the [CRAN website](https://cran.r-project.org/) and is available for all operating systems.  Updated versions of R are released frequently, and it is recommended that you have the latest version installed. 

Head to the [CRAN website](https://cran.r-project.org/) to download the latest version of R. 

You can check your version of R by typing `version` into the R console.

### Installing RStudio  
[RStudio](https://www.rstudio.com/products/RStudio/) is a freely available integrated development environment (IDE) for R.  It is not essential that you use RStudio, but it can make things a lot easier, especially if you are new to R. 

We assume throughout the `openMSE` documentation that you are using RStudio.  

It is important to be aware that RStudio and R are two different pieces of software that must be installed separately.  We recommend installing the R software before downloading and installing RStudio. 

### Installing `openMSE`

We recommend installing `openMSE` using [`pak`](https://pak.r-lib.org/reference/pak.html), which automatically checks that all dependency packages meet the minimum version requirements:

```{r}
#| eval: false
# Install pak if required
if (!requireNamespace("pak", quietly = TRUE))
  install.packages("pak")

# Install openMSE and dependencies
pak::pak("openMSE")
```

To upgrade all packages to the latest versions:

```{r}
#| eval: false
pak::pak("openMSE", upgrade = TRUE)
```

Then load `openMSE` into your R session:

```{r}
#| eval: false
library(openMSE)
```

::: callout-important
## Development Version

`MSEtool v4.0` has not yet been released on CRAN and must be installed from GitHub:

```{r}
#| eval: false
pak::pkg_install("blue-matter/MSEtool")
```

All code in this manual works with this development version. The existing CRAN versions of `openMSE`, `DLMtool`, and `SAMtool` are compatible. This note will be removed once `MSEtool v4.0` is released on CRAN.
:::

## About Version 2.0

`openMSE 2.0` is a major update featuring a complete redesign of the core S4 object classes and a full refactoring of the source code. It is backwards compatible; code written for earlier versions should continue to work without modification, though for new work we recommend following the methods described in this manual. For a full description of changes see @sec-changes.

