# Built-In Examples {#sec-buildexamples}

```{r}
#| echo: false
#| include: false
library(MSEtool)
rnorm(5)
```

This chapter describes the built-in example objects included with `openMSE` and walks through how each is constructed.

The objects are built manually here, with the parameters are specified directly by the user. This is the most transparent approach and the easiest way to understand the structure of each object and its subcomponents.

However, this is not the only way to construct operating models. `openMSE` also supports importing parameter estimates directly from stock assessment output (e.g., Stock Synthesis), as well as conditioning an operating model on observed fishery data such as catch histories and abundance indices. These alternative approaches are covered in other chapters in this section.

## Example `Stock` Objects {#sec-examplestocks}

A `Stock` object defines the biological and demographic characteristics of a fish population in the operating model. It is composed of a set of subcomponent objects including age structure, length, weight, natural mortality, and maturity schedules, the stock-recruitment relationship, and spatial dynamics.

See `?Stock` for full documentation of the `Stock` object and its subcomponents.

Three example `Stock` objects are provided, representing species with contrasting life histories and model configurations:

- **`AlbacoreExStock`**: a relatively long-lived, slow-growing large pelagic species;
- **`ButterfishExStock`**: a short-lived, fast-growing forage species;
- **`SeasonalSpatialExStock`**: a short-lived reef fish with monthly time steps, seasonal recruitment, and age-based movement across three areas.

### `AlbacoreExStock`

```{r}
#| include: false
knitr::read_chunk("../MSEtool/data-raw/ex-stock-albacore.R")
```

`AlbacoreExStock` represents an Albacore tuna (*Thunnus alalunga*) stock, a relatively long-lived, slow-growing large pelagic species with low natural mortality, moderate steepness, and relatively low recruitment variability.

#### Initialize the Stock

The `Stock` object is created with a `Name`, which serves as a unique identifier, along with the optional `CommonName` and `Species` metadata fields for display and documentation purposes. The object acts as a container for the biological and demographic subcomponents assigned in the steps that follow.

```{r albacore-create-stock}
```

#### Age Structure

The age structure of the stock is defined by setting the maximum age (assumed to be a plus-group by default).

The `Ages` function is used to initialize an `ages-class` object and assign it to the `AlbacoreExStock` stock object created in the previous step.

See `?Ages` for more information on the `Ages` object.

```{r albacore-create-ages}
```

#### Length-at-Age

Growth for this stock follows a von Bertalanffy model, with `Linf`, `K`, and `t0` each specified as uniform ranges.

The `CVatAge` parameter captures individual variation in length at age around the mean growth trajectory.

As with `Ages` above, the `Length` function is used to create the `length-class` object and assign it to the `AlbacoreExStock` stock object.

The remaining subcomponents follow the same pattern: each is constructed using its corresponding constructor function and assigned to the `Stock` object via the matching slot-assignment operator (see @sec-assignmentfunctions).

See `?Length` for more information on the `Length` object and `?LengthModels` for other length-at-age models included in `openMSE`.

```{r albacore-create-length}
```

#### Weight-at-Age

An allometric length-weight relationship $W = \alpha L^\beta$ is used to convert mean length-at-age to a weight-at-age schedule.

In this example the parameters are specified as fixed point values rather than ranges.

See `?Weight` for more details on the `Weight` object.

```{r albacore-create-weight}
```

#### Natural Mortality

Natural mortality $M$ is specified as a uniform prior range, reflecting uncertainty in this parameter.

In this example, $M$ is assumed age-invariant and constant across time.

See `?NaturalMortality` and `?NaturalMortalityModels` for more details on the `NaturalMortality` object and supported models.

```{r albacore-create-natural-mortality}
```

#### Maturity

The maturity ogive is described by a logistic function parameterized by $L_{50}$ (length at 50% maturity) and $L_{50\_95}$ (the length increment from 50% to 95% maturity), each specified as ranges of a uniform distribution.

Spawning biomass is calculated by multiplying the maturity- and weight-at-age schedules. By default, spawning production is assumed proportional to spawning biomass.

Users can optionally specify a `Fecundity` object that defines egg production at age explicitly; in this case, spawning production is calculated directly from the `Fecundity` schedule, and the `Maturity` object is used only to identify mature fish rather than to scale reproductive output.

See `?Maturity`, `?MaturityModels`, and `?Fecundity` for full documentation.

```{r albacore-create-maturity}
```

#### Stock-Recruitment Relationship

A Beverton-Holt stock-recruitment relationship is used for this stock. It is parameterized by steepness $h$, which defines the fraction of unfished recruitment produced at 20% of unfished spawning biomass.

`R0` sets the unfished equilibrium recruitment in numbers, establishing the scaling of the population. Process error in recruitment is modelled as log-normal deviations with standard deviation `SD` and first-order autocorrelation `AC`.

In this example, all parameters except `R0` are defined as the bounds of a uniform distribution.

See `?SRR` and `?SRRModels` for more details on the `SRR` object and supported models.

```{r albacore-create-srr}
```

#### Spatial Structure

The `Albacore` stock is distributed across two areas in this example.

`UnfishedDist` sets the lower and upper bounds of a uniform distribution for the equilibrium proportion of unfished biomass in Area 1; `ProbStaying` is the annual probability that an individual remains in Area 1; and `RelativeSize` is the relative habitat size of Area 1. The complement of each quantity applies to Area 2.

The `Spatial` subcomponent is optional. If omitted, the model assumes a single well-mixed area. When spatial structure is included, all stocks within an operating model must share the same spatial configuration (i.e., the same number of areas). `openMSE` supports an unlimited number of areas and additional features such as age-based movement rates. See `?Spatial` for full documentation.

```{r albacore-create-spatial}
```

#### Depletion {#sec-exampledepletion}

The `Depletion` subcomponent is optional.

When specified, it defines the stock's depletion level at the beginning and/or end of the historical period, expressed as a fraction of unfished biomass $\frac{B}{B_0}$ (other definitions of depletion can be specified using the `Reference` argument).

In this example, terminal depletion is given a wide range of 5–60% of $B_0$, reflecting broad uncertainty in current stock status.

When `Final` is specified, the operating model adjusts each fleet's catchability so that the simulated terminal depletion matches the specified distribution (see @sec-examplefleets). This ensures that each `Stock` object remains compatible with any `Fleet` object, regardless of the effort history or catchability values the fleet specifies.

In practice, users will typically parameterize fleets with an explicit effort time series and catchability values that reflect the history of the fishery as estimated by a stock assessment or conditioning model. In these cases, `Final` should be left unspecified and terminal depletion will be determined by the interaction between effort, catchability, and the population dynamics.

Specifying `Final` alongside explicit catchability values is not recommended, as the optimization will override the user-specified catchability and may produce unexpected results.

See `?Depletion` for full documentation.

```{r albacore-create-depletion}
```

### `ButterfishExStock`

```{r}
#| include: false
knitr::read_chunk("../MSEtool/data-raw/ex-stock-butterfish.R")
```

`ButterfishExStock` represents a Butterfish (*Peprilus triacanthus*) stock, a short-lived, fast-growing forage species with high natural mortality and wide uncertainty on recruitment and maturity.

#### Initialize the Stock

```{r butterfish-create-stock}
```

#### Age Structure

A maximum age of 8 years is assumed for this stock. As with the Albacore example described above, this maximum age class is used as a plus-group.

```{r butterfish-create-ages}
```

#### Length-at-Age

Length-at-age is again modeled wit the `vonBert` model.

```{r butterfish-create-length}
```

#### Weight-at-Age

The same allometric length-weight relationship is specified with fixed point values.

```{r butterfish-create-weight}
```

#### Natural Mortality

Natural mortality is substantially higher than for Albacore (0.7–0.9), again assumed constant across ages and timesteps.

```{r butterfish-create-natural-mortality}
```

#### Maturity

The same logistic parameterization is used as for Albacore, but with substantially wider prior ranges on both $L_{50}$ and $L_{50\_95}$, reflecting greater uncertainty in the maturity schedule of this species.

```{r butterfish-create-maturity}
```

#### Stock-Recruitment Relationship

Recruitment variability is considerably higher than for Albacore (`SD` of 0.7–1.1 versus 0.15–0.3), consistent with the volatile year-class dynamics typical of short-lived forage species.

```{r butterfish-create-srr}
```

#### Spatial Structure

The lower `ProbStaying` values (0.4–0.6) relative to Albacore reflect greater mixing between areas.

```{r butterfish-create-spatial}
```

#### Depletion

The same broad prior range as `AlbacoreExStock` is used, for the same reasons described above.

```{r butterfish-create-depletion}
```

### `SeasonalSpatialExStock`

```{r}
#| include: false
knitr::read_chunk("../MSEtool/data-raw/ex-stock-seasonalspatial.R")
```

`SeasonalSpatialExStock` is a made-up example designed to demonstrate two features not present in the other example stocks: 

1. seasonal recruitment patterns;
2. age-based movement across spatial areas. 

Unlike the previous two examples, this stock is built in a single `Stock()` constructor call, with all subcomponents passed directly as arguments. This is an alternative to the step-by-step accessor assignment style used above; both approaches produce identical objects. 

The only exception is the spatial distribution arrays, which in this example are constructed separately before being passed to `Spatial()`.

All parameters are fixed single values rather than uniform prior ranges in this example, so the object produces identical dynamics across all simulations (with the exception of the stochastic recruitment deviations).

#### Spatial Distribution Arrays

Three areas represent a linear inshore-to-offshore gradient. The unfished biomass distribution is described by Gaussian curves centred at the start, middle, and end of the lifespan, so that juveniles concentrate in Area 1 (e.g., an inshore lagoon, 15% of habitat), subadults in Area 2 (e.g., mid-shelf reef, 35%), and adults in Area 3 (e.g., outer slope, 50%). 

`FracOther` connects adjacent area pairs freely but provides negligible connectivity between Areas 1 and 3.

```{r seasonalspatial-create-spatial-arrays}
```

#### Build the Stock

With the spatial arrays already constructed, the complete `Stock` object is generated in a single call. The key features to note are:

- `Ages(MaxAge = 36, Units = 'month')`: age classes run from 0 to 3 years in monthly steps. This stock must be paired with an `OM` with `Seasons = 12`.
- `NaturalMortality(Pars = list(M = 0.09))`: $M$ is 0.09 *per month* , consistent with ≈1–5% survival to `MaxAge`. Because `Ages@Units = 'month'`, $M$ must be expressed on the same per-time-step basis (see @sec-populate).
- `SetSeasonalR0()`: distributes annual unfished recruitment across months as a Gaussian pulse peaking in month 6, creating a sharp seasonal spawning pattern.
- `Spatial()`: the pre-computed `UnfishedDist` and `FracOther` arrays are passed directly, along with area-specific `ProbStaying` and `RelativeSize` values.

```{r seasonalspatial-create-stock}
```

## Example `Fleet` Objects {#sec-examplefleets}

A `Fleet` object defines the exploitation characteristics of a fishing fleet in the operating model. It is composed of subcomponent objects including a historical effort trajectory, selectivity-at-age or -at-length, and optional components such as catchability, retention, and discard mortality.

See `?Fleet` for full documentation of the `Fleet` object and its subcomponents.

Two example `Fleet` objects are provided, representing gear types with contrasting selectivity patterns:

- **`AsympExFleet`**: a fleet with asymptotic (logistic) selectivity, retaining all fish above the fully-selected length;
- **`DomeExFleet`**: a fleet with dome-shaped selectivity, with reduced retention of the largest fish.

Both fleets use the same general effort trajectory, rising through the historical period and stabilising at the terminal year, but differ in both the rate of that rise and in their selectivity curves.

### `AsympExFleet`

```{r}
#| include: false
knitr::read_chunk("../MSEtool/data-raw/ex-fleet-asymp.R")
```

`AsympExFleet` represents a fleet using gear with asymptotic selectivity.

#### Initialize the Fleet

The `Fleet` object is created with a `Name` that serves as a unique identifier. As with `Stock`, it acts as a container for the subcomponent objects assigned below.

```{r asymp-create-fleet}
```

#### Effort

In this example, historical effort is specified as a data frame of control points, from which `nSim` stochastic effort trajectories are generated by the operating model.

Each control point defines `Year` (as a normalized index from 0 to 1 spanning the historical period), and `Lower` and `Upper` bounds of a uniform distribution from which relative effort at that point is drawn independently for each simulation. The resulting trajectories are linearly interpolated between control points, with multiplicative lognormal process error with coefficient of variation `CV`, and normalized so that effort equals 1 in the terminal year.

In this example, effort starts at zero, rises to a plateau between 40% and 60% of maximum by the midpoint of the historical period, and remains at full effort through to the terminal year.

This control-point approach is typical of data-limited settings where absolute effort records are unavailable. It allows users to sketch plausible trends in fishing pressure, e.g., a gradual build-up, a period of stability, then a decline, without requiring a precise time series. @fig-asymp-effort shows the resulting stochastic trajectories generated from the control points defined above.

Because the resulting effort index is relative and normalized to 1 at the terminal year, it carries no absolute scale; catchability is then back-calculated from the value of `Final` in the `Depletion` component of the corresponding stock object (see @sec-exampledepletion).

In practice, most users would instead supply an explicit effort time series derived from stock assessment output (often assumed proportional to fishing mortality), paired with a `Catchability` object, and leave `Depletion@Final` unspecified.

See `?Effort` and `?GenHistEffort` for full documentation of the `Effort` object and the control-point data frame format.

```{r asymp-stochastic-effort}
```

The following code block demonstrates the resulting effort trajectories specified above:

```{r  fig-asymp-effort}
#| fig-cap: "Ten stochastic historical effort trajectories for `AsympExFleet`, generated from the control-point data frame. Effort is relative and normalized to 1 in the terminal year; the spread across trajectories reflects uncertainty in the level of effort at the intermediate control points, with additional lognormal process error (CV = 0.1) applied within each simulation."
```

#### Selectivity

Selectivity is specified using a double-normal length-based model parameterized by three quantities:

- `L5`, the length at 5% selectivity;
- `LFS`, the length at full (100%) selectivity; and
- `Vmaxlen`, the selectivity of the longest fish in the population.

Setting `Vmaxlen = 1` produces a purely asymptotic curve, selectivity rises from `L5` to `LFS` and remains at 1 for all larger fish (@fig-asymp-selectivity).

With `isRel = TRUE`, the `L5` and `LFS` values are interpreted as multiples of $L_{50}$ (the length at 50% maturity) of the paired stock rather than as absolute lengths. This is used here so that example fleets can be combined with the example stocks without modification.

In practice, users would typically specify selectivity in absolute length units tailored to the specific stock being modelled. The pairing between a fleet and a stock occurs when the fleet is added to an `OM` object; see @sec-exampleOM for details.

See `?Selectivity` and `?SelectivityModels` for full documentation and available model families.

```{r asymp-selectivity}

```

The code block below demonstrates the resulting selectivity schedules when this `Selectivity` object is paird with the `AlbacoreExStock` stock object.

It first manually populates the `Maturity` and `Selectivity` objects for `AlbacoreExStock` and `AsympExFleet` respectively, then plots the resulting at-age and at-length schedules. Populating these objects expands the parameter inputs (the bounds vectors specified in the constructor calls above) into the full `Sim × Age × Year` and `Sim × Class × Year` arrays, sampling one value per simulation (here `nSim = 5`) from each uniform range.

In normal use this happens automatically when `Populate()` is called on the `OM` object; it is done explicitly here for illustration only. See @sec-populate for a full description of the population process.

```{r fig-asymp-selectivity}
#| fig-cap: "Selectivity (blue) and maturity (coral) schedules for `AsympExFleet` paired with `AlbacoreExStock`, shown at length (left) and age (right). Each line represents one simulation replicate; the spread reflects uncertainty in the growth and maturity parameters sampled across simulations. Length-based parameters (`L5`, `LFS`) are specified relative to $L_{50}$ of the paired stock (`isRel = TRUE`), so the position of the selectivity curve varies across simulations in proportion to the sampled maturity schedule."
```

#### Optional Subcomponents

`Retention`, `DiscardMortality`, and `Catchability` are not specified for `AsympExFleet`. The first two are optional: when omitted, the model assumes all age and size classes are fully retained and that discarded fish have zero mortality. See `?Retention` and `?DiscardMortality` for full documentation.

`Catchability` is a special case. It is left empty here because, as described above, the example stocks have `Depletion@Final` specified, which instructs the operating model to calculate the catchability required to achieve the target terminal depletion for each simulation (see the `Depletion` subsection under each example stock above).

In practice, users would typically specify catchability explicitly based on stock assessment estimates and leave `Depletion@Final` unspecified, allowing terminal depletion to emerge from the interaction between effort, catchability, and population dynamics.

### `DomeExFleet`

```{r}
#| include: false
knitr::read_chunk("../MSEtool/data-raw/ex-fleet-dome.R")
```

`DomeExFleet` represents a fleet with dome-shaped selectivity. Selectivity rises through the ascending limb in the same way as `AsympExFleet`, but declines for the largest fish.

#### Initialize the Fleet

```{r dome-create-fleet}
```

#### Effort

The effort trajectory for `DomeExFleet` follows the same control-point structure as `AsympExFleet`, but the mid-period plateau is lower (0.4–0.6 of maximum), indicating a slower build-up of fishing pressure before the final ramp-up to full effort at the terminal year.

```{r dome-stochastic effort}
```

#### Selectivity

The dome-shaped curve is produced by setting the `Vmaxlen` parameter of the `DoubleNormal` model to a range of 0.5–1 rather than a fixed value of 1.

`Vmaxlen` is the selectivity at the largest length class (defined in `Classes`). Values below 1 reduce selectivity at the largest sizes, with `Vmaxlen = 0.5` meaning that fish in the largest length class are only half as likely to be caught as fish at the fully-selected length (@fig-dome-selectivity).

See `?SelectivityModels` for more details on `DoubleNormal` and other supported selectivity models.

```{r dome-selectivity}
```

```{r fig-dome-selectivity}
#| echo: false
#| fig-cap: "Selectivity (blue) and maturity (coral) schedules for `DomeExFleet` paired with `AlbacoreExStock`, shown at length (left) and age (right). Each line represents one simulation replicate; the spread reflects uncertainty in the growth and maturity parameters sampled across simulations. Length-based parameters (`L5`, `LFS`) are specified relative to $L_{50}$ of the paired stock (`isRel = TRUE`), so the position of the selectivity curve varies across simulations in proportion to the sampled maturity schedule."
```

## Example `Obs` Objects {#sec-exampleobs}

An `Obs` object defines the observation model for a fleet and its corresponding stock or stock complex. It describes the structure of sampling error and systematic bias for each data type. Each fleet object has its own `Obs` object.

See `?Obs` for full documentation. See `?CatchObs`, `?IndicesObs`, and `?CompObs` for documentation of the subcomponent constructors.

The `Obs` object interacts closely with the corresponding `Data` object. See @sec-dataobs for more details.

Five example `Obs` objects are provided, representing monitoring programmes of increasing data richness:

- **`CommercialFleetObs`**: landings, discards, effort, and a CPUE index from a commercial fleet.
- **`CatchAndSurveyObs`**: landings and a fishery-independent survey index only.
- **`AgeStructuredObs`**: landings, discards, a survey index, and age composition samples from both landings and discards.
- **`LengthStructuredObs`**: landings, discards, a survey index, and length composition samples from both landings and discards.
- **`DataRichObs`**: all data types ares generated: landings, discards, effort, CPUE, survey, and both age and length compositions for landings and discards.

### Observation Error Parameters

Before describing the individual objects, it is useful to understand the parameters common to all observation subcomponents.

**`CatchObs`** (used for `Landings` and `Discards`) takes two main parameters:

- `CV`: coefficient of variation of lognormal observation error on catch. Accepts a scalar (same CV for all simulations), a length-2 vector `c(lower, upper)` (a unique CV per simulation is drawn from `Uniform(lower, upper)`), or a length-`nSim` array (used directly). After population, `@CV` is a named `[nSim]` vector and `@Error` becomes a named `[nSim × nYear]` lognormal array.

- `Bias`: multiplicative observation bias on the natural scale. `NULL` defaults to 1 (unbiased). A length-`nSim` vector is used directly as one bias per simulation. Any other scalar input is treated as the CV of a lognormal from which `nSim` bias values are drawn — e.g. `Bias = 0.1` draws biases centred at 1 with CV of 10%. After population, `@Bias` is a named `[nSim]` array of positive multipliers.

**`EffortObs`** (used for `Effort`) takes the same `CV` and `Bias` parameters as `CatchObs`.

**`IndicesObs`** (used for `CPUE` and `Survey`) takes:

- `CV`: coefficient of variation of lognormal index observation error.
- `AC`: lag-1 autocorrelation of index residuals, reflecting persistent deviations in survey catchability or other sources of correlated index error.
- `Selectivity`: how the simulated population is aggregated to produce the index. `"Biomass"` applies flat selectivity (total biomass); `"SBiomass"` uses maturity-at-age as selectivity (equivalent to a spawning biomass index). For CPUE indices, the corresponding fleet selectivity is used automatically when `Selectivity` is `NULL`.

**`CompObs`** (used for all composition slots) takes:

- `SampleSize`: nominal number of fish aged or measured per year. A length-2 vector draws a unique value per simulation from `Uniform(lower, upper)`. Setting `SampleSize = NULL` suppresses composition data generation entirely.
- `ESS`: effective sample size, which controls stochastic variability in the composition draw independently of `SampleSize`. Values of `ESS < SampleSize` produce overdispersion relative to a pure multinomial — typical of age or length compositions that contain correlated sampling error.
- `Theta`: Dirichlet-Multinomial dispersion parameter in $(0, 1]$. `Theta = 1` recovers a standard multinomial draw; values approaching 0 produce maximum overdispersion.

### `CommercialFleetObs`

```{r}
#| include: false
knitr::read_chunk("../MSEtool/data-raw/ex-obs-comm-fleet.R")
```

`CommercialFleetObs` represents a commercial fleet monitoring programme that produces records of total landings & discards, fishing effort, and an index of vulnerable abundance. Landings and discards are observed with moderate error and some positive bias in landings (reflecting potential over-reporting or high-grading), effort is recorded with low error, and a CPUE index is available.

::: callout-note
## CPUE and Survey Indices

Note that CPUE (and survey indices) in `openMSE` are calculated directly from the OM population abundance, not derived from the ratio of observed catch to observed effort. This is equivalent to assuming the index has been standardised so that it reflects true abundance rather than the conflated product of catchability, effort, and catch reporting error.

The index is assumed proportional to abundance (i.e. a constant catchability coefficient), with no hyperdepletion or hyperstability. The `CV` and `AC` parameters in `IndicesObs` represent the residual observation error around that proportional relationship after standardisation.
:::

```{r comm-create-obs}
```

Landings CV of 5–15% and a bias range of 1.0–1.15 reflect moderately reliable catch reporting with a tendency toward slight overestimation. Discard CV of 20–40% and a wide bias range of 0.8–1.2 reflect the greater uncertainty typical of discard estimation. The CPUE index has CV of 15–25% and moderate autocorrelation (AC 0–0.3).

### `CatchAndSurveyObs`

```{r}
#| include: false
knitr::read_chunk("../MSEtool/data-raw/ex-obs-catch-survey.R")
```

`CatchAndSurveyObs` represents a minimal data scenario: only landed catch and a fishery-independent survey index are available.

```{r catch-survey-create-obs}
```

The survey uses `Selectivity = "Biomass"`, meaning it is treated as a total biomass index with flat selectivity across ages. Landings CV of 10–25% and a downward bias range of 0.8–1.0 reflect the possibility of underreporting.

### `AgeStructuredObs`

```{r}
#| include: false
knitr::read_chunk("../MSEtool/data-raw/ex-obs-age-structured.R")
```

`AgeStructuredObs` represents a data-moderate programme with age-structured biological sampling. In addition to landings and discards, a fishery-independent survey and age composition samples from both the landed and discarded catch are available.

```{r age-structured-create-obs}
```

The survey uses `Selectivity = "SBiomass"`, treating it as a spawning biomass index. Age compositions are sampled with `SampleSize` of 200–400 (landings) and 100–200 (discards), with effective sample sizes of 50–100 and 30–80 respectively. `Theta` values of 0.4–1.0 allow for additional Dirichlet-Multinomial overdispersion.

### `LengthStructuredObs`

```{r}
#| include: false
knitr::read_chunk("../MSEtool/data-raw/ex-obs-length-structured.R")
```

`LengthStructuredObs` is structured analogously to `AgeStructuredObs`, but with length composition samples in place of age compositions. This reflects monitoring programmes where length measurement is feasible but ageing is not.

```{r length-structured-create-obs}
```

Length composition sample sizes are somewhat larger than the age-structured equivalents (300–600 for landings), consistent with the lower cost of length measurement relative to ageing.

### `DataRichObs`

```{r}
#| include: false
knitr::read_chunk("../MSEtool/data-raw/ex-obs-data-rich.R")
```

`DataRichObs` represents a comprehensive data-rich monitoring programme, combining all available data types: landings, discards, effort, a CPUE index, a fishery-independent survey, and both age and length compositions for landings and discards.

```{r data-rich-create-obs}
```

Observation errors throughout are tighter than in the other example objects: landings CV of 3–8%, effort CV of 2–5%, and survey CV of 8–15% with very low autocorrelation (AC 0–0.1). The survey uses `Selectivity = "SBiomass"` as an index of spawning biomass, while the CPUE index uses fleet selectivity automatically. Composition sample sizes are the largest of all example objects, with ESS values approaching the nominal sample sizes, reflecting high-quality age and length data.

## Example `Imp` Objects

::: callout-note
The `Imp` class is currently a placeholder and is not yet implemented.

This section will be updated once example `Imp` objects are developed.
:::

## Example `OM` Objects {#sec-exampleOM}

Five example `OM` objects are provided, covering the most common configurations of stocks, fleets, and observation programmes. All use the example `Stock` and `Fleet` objects described in @sec-examplestocks and @sec-examplefleets, and the example `Obs` objects described in @sec-exampleobs.

Each object is initialised with 8 stochastic simulations (`nSim = 8)`, a 20-year historical period (`nYear = 20)`, and a 30-year projection period (`pYear = 30`).

### `SingleStockOM`

```{r}
#| include: false
knitr::read_chunk("../MSEtool/data-raw/ex-om-singlestock.R")
```

The simplest configuration: one stock and one fleet.

`AlbacoreExStock` is fished by `AsympExFleet` with age-structured data collection (`AgeStructuredObs`).

```{r singlestock-create-om}
```

When a single `Stock`, `Fleet`, or `Obs` object is assigned without a list wrapper, `openMSE` treats it as a single-stock, single-fleet model.

Alternatively, the `Stock`, `Fleet`, and `Obs` objects could also be assigned with assignment functions, either directly as the objects:

```{r singlestock-assignment}
#| eval: false
```

or explictly as lists and nested lists that represent the final structure:

```{r singlestock-assignment2}

```

### `TwoFleetOM`

```{r}
#| include: false
knitr::read_chunk("../MSEtool/data-raw/ex-om-two-fleet.R")
```

One stock fished by two fleets with different selectivity profiles and data collection programs.

`AlbacoreExStock` is fished by `AsympExFleet` (with `AgeStructuredObs`) and `DomeExFleet` (with `CatchAndSurveyObs`), representing a scenario where the dome-selected fleet has only catch and survey data available.

```{r two-fleet-create-om}
```

For a single stock with multiple fleets, `Fleet` and `Obs` are supplied as nested lists with one outer element per stock and one inner element per fleet: `Fleet[[stock]][[fleet]]`. The outer list must always be present even for a single stock.

### `MultiStockOM`

```{r}
#| include: false
knitr::read_chunk("../MSEtool/data-raw/ex-om-multistock.R")
```

Two stocks each fished by two fleets, with different observation programmes per stock-fleet combination.

`AlbacoreExStock` is fished by both `AsympExFleet` (with `AgeStructuredObs`) and `DomeExFleet` (with `CommercialFleetObs`); `ButterfishExStock` is fished by the same two fleets but monitored with `CatchAndSurveyObs` and `LengthStructuredObs` respectively.

```{r multistock-create-om}
```

`Fleet` and `Obs` are both indexed `[[stock]][[fleet]]`. All stocks must have the same number of fleets. 

In this example both stocks share the same fleet objects. However, in practice, fleet specifications typically differ between stocks even when the same physical gear is operating on both populations, since selectivity, retention, discard
mortality, and effort can all vary by stock.

### `ComplexOM`

```{r}
#| include: false
knitr::read_chunk("../MSEtool/data-raw/ex-om-complex.R")
```

Two stocks managed as a single stock complex, fished by one fleet with one shared observation programme. 

The `Complexes` argument groups both stocks into a single complex named `StockComplex`, meaning fishery data are reported, and management advice is implemented, at the complex level rather than per stock.

```{r complex-create-om}
```

When stocks are grouped into complexes, `Obs` must be indexed by complex rather than by stock (`length(Obs) == nComplex`). Here a single `AgeStructuredObs` object is assigned and applied to the whole complex. The fleet operates on both stocks simultaneously; the relative contribution of each stock to the complex catch is determined by their relative abundance and the shared selectivity schedule.

As with `MultiStockOM` above, in practice the fleet properties would typically vary over stocks.

See @sec-complexes for more details on stock complexes.

### `SeasonalSpatialOM`

```{r}
#| include: false
knitr::read_chunk("../MSEtool/data-raw/ex-om-seasonalspatial.R")
```

`SeasonalSpatialOM` pairs `SeasonalSpatialExStock` with `DomeExFleet` and `CommercialFleetObs`. It is the only example OM with sub-annual time steps (`Seasons = 12`) and a multi-area spatial model. In all other respects it follows the same single-stock, single-fleet structure as `SingleStockOM`.

```{r seasonalspatial-create-om}
```

