# The Stock Object {#sec-stock}

```{r,echo=FALSE}
suppressMessages(library(MSEtool))
```

`Stock` objects are a sub-component of the [Operating Model Object](\operatingmodel.html).

The `Stock` object contains the biological parameters for a modeled population.

`Stock` objects are created with the `Stock()` function. The first three arguments to the `Stock` function are used to set a `Name` for the `Stock` object, and a `CommonName` and `Species` name (genus and species) for the stock being modeled in the operating model.

For example:

```{r}
Albacore <- Stock(Name='Albacore Stock',
                       CommonName='Albacore',
                       Species='Thunnus alalunga')
```

The various biological characteristics of the stock, including growth, maturation, natural mortality, stock-recruitment relationship, and spatial distribution and movement rates, are stored in sub-objects within the `Stock` object.

The sections below include some examples for populating these sub-objects for the example `Albacore` stock created above. Note that these examples are very demonstrations purposes and are not intended to reflect real values for an Albacore stock (i.e., the values are made up!).

For more information on the `Stock` object, see the help documentation with `?Stock`.

## Ages

The `Ages` object contains information relating to the age classes for the stock, specifically:

-   `MaxAge`: `r MSEtool:::MaxAge_param()`

-   `Units`: `r MSEtool:::Units_param()`

-   `PlusGroup`: `r MSEtool:::Plusgroup_param()`

`Ages` objects are created and populated in the usual way, and added to
the `Stock` object:

```{r}
Ages(Albacore) <- Ages(MaxAge=20)
```

Alternatively, the `Ages` object (and all other sub-objects in the `Stock` object) can be added directly in the call to the `Stock()` function:

```{r}
AlbacoreStock <- Stock('Albacore Stock',
                       'Albacore',
                       'Thunnus alalunga',
                       Ages=Ages(20))
```

## Length

The `Length` object contains information relating to mean length-at-age and the distribution around the mean length-at-age. 

Although most `Stock` objects will require a `Length` object, technically the `Length` object is optional and can be left blank if it is not needed. 

The information in the `Length` object is used within the operating model for two purposes:

1. Generate the required [maturity-](stockobject.html#maturity) or [selectivity-at-age](fleetobject.html#selectivity) schedules from maturity- or selectivity-at-length models;
2. Simulate catch-at-length data to be used within [Management Procedures](/aboutMPs.html).

If catch-at-length data is not required and the maturity- and selectivity-at-age schedules are provided directly (i.e., not via a length-based model or schedule), then the `Length` object is not required in the operating model.

The mean length-at-age information can be entered in two ways: 

1. With a growth model and associated parameters;
2. Passed directly as a 3-dimensional array;.

### Specifying Length-at-Age Model with `Pars` and `Model`

The `Pars` slot in the `Length` object is used to store the parameters of a model that generates the mean length-at-age schedule.

```{r, include=FALSE}
models <- LengthModels(print=FALSE)
nmodels <- nrow(models)
txt <- ifelse(nmodels>1, 'are', 'is')
```

Currently there `r txt` `r nmodels` length models built into `MSEtool`. The `LengthModels()` function can be used to display the built-in growth models in the R console.

To use a built-in model, `Pars` must be a named list, with the names matching those listed under the `Pars` displayed by `LengthModels()` for the relevant model.

For example, to use the von Bertalanffy growth model (`vonBert`):

```{r}
length <- Length(Pars=list(Linf=100, 
                           K=0.2,
                           t0=-0.1))
```

`Model` will be automatically populated with the matching model name:

```{r}
Model(length)
```

An error will occur if the named elements of `Pars` does not exactly match the parameters in one of the models listed in `LengthModels()`, e.g.,

```{r}
#| error=TRUE
length <- Length(Pars=list(Linf=100, 
                           K=0.2))
```

#### Structure of `Pars`

The elements in the `Pars` list can be structured in several different ways to deal with 




```{r}
LengthModels()
```


### Entering Length-at-Age Directly with `MeanAtAge`

Alternatively, the mean length-at-age schedule can be passed as a 3-dimensional array directly into the `MeanAtAge` slot:


```{r}


length <- Length(MeanAtAge = )
```




### Distribution of Length-at-Age

### Units 

`Units`

### Timing

### Classes 

## Weight

## Natural Mortality

## Maturity

## Fecundity

## Stock-Recruit Relationship

## Spatial

## Depletion

## Misc 
