```{r}
#| eval: true
#| include: false
#| message: false
#| warning: false
#| cache: false
library(MSEtool)
knitr::read_chunk('../MSEtool/data-raw/ExampleOM.R')
```

# Build an `OM` Object {#sec-om_build}

## Create a new `OM` Object

The `OM()` function is used to create a new Operating Model object (see `?OM`). The contents of the `OM` can be specified with arguments to the `OM` function:

```{r initialize-om }

```

or by using functions matching the names of the slots in the object.

For example, here we'll change the number of independent stochastic simulations in our OM from the default `nSim=48` to a small number so that the code runs very quickly:

```{r access-assign-1}
```

Similarly, the number of historical years and projection years can be changed:

```{r access-assign-2}

```

The contents of the `OM` object, and the various methods to access and modify these contents are discussed in more detail in @sec-object_om.

## Populating the `OM` Object

An `OM` object needs at least one [`Stock`](object_stock.html) and one [`Fleet`](object_fleet.html) object before it can be used to `Simulate` a historical fishery and `Project` it forward for the closed-loop simulation testing (@fig-flowchart).

The process to build these objects is described in the next few chapters. In @sec-om_populate, these objects are added to the new `ExampleOM` object created here.
